---
title_meta  : บทที่ 4
title       : Factors
description : มีอยู่บ่อยครั้งที่ข้อมูลของเราถูกเก็บอยู่ในรูปของข้อมูลที่ใช้แบ่งกลุ่ม (categorical data) ยกตัวอย่างเช่น มนุษย์นั้นมี 2 เพศคือชายและหญิง ใน R ข้อมูลที่แสดงถึงการแบ่งกลุ่มจะถูกเก็บในรูปแบบที่เรียกว่า factor ซึ่งเราจะให้คุณได้เรียนรู้ถึงวิธีการสร้างและจัดการกับข้อมูลประเภทนี้ในบทนี้

--- type:NormalExercise xp:100 skills:1 key:05273321916d99bb9c0deadf75c6834d25a47244
## factor คืออะไรและทำไมเราถึงต้องใช้มัน?

ในบทนี้ เราจะให้คุณได้เรียนรู้เกี่ยวกับข้อมูลอีกแบบหนึ่งซึ่งเรียกว่า **factor**

คำว่า factor นั้นอ้างอิงถึงข้อมูลในทางสถิติที่ใช้ในการแบ่งกลุ่มข้อมูล ข้อแตกต่างระหว่างข้อมูลที่ใช้แบ่งกลุ่ม(categorical data) กับข้อมูลอื่นๆทั่วไปหรือแบบต่อเนื่อง (continuous data) คือ ข้อมูลที่ใช้แบ่งกลุ่มนั้นจะใช้**ข้อมูลเพียงไม่กี่ตัว**(เช่น เพศชาย เพศหญิง) ในขณะที่ข้อมูลอื่นๆทั่วไปสามารถมีค่าได้หลากหลายกว่าจนอาจจะไม่สามารถนับได้(เช่น เลข 1 ถึง 1,000)

การที่เราระบุประเภทข้อมูลให้ R รู้อย่างชัดเจนว่า ข้อมูลนี้คือ categorical หรือ continuous จะช่วยให้ R สามารถทำงานได้ดีขึ้นในการสร้าง model ทางสถิติซึ่งสามารถนำมาใช้วิเคราะห์ข้อมูบได้ในอนาคต (เนื่องจาก R จะปฏิบัติต่อ categorical data แตกต่างจาก continuous data)

ตัวอย่างที่ดีของตัวแปรที่ใช้แบ่งกลุ่มคือ 'เพศ' เนื่องจากมนุษย์ทุกคนสามารถถูกแบ่งเป็นกลุ่มได้ตามเพศแค่ "ชาย" หรือ "หญิง" เท่านั้น เราจึงสามารถใช้ 'เพศ' เป็นตัวแปรในการแบ่งกลุ่มได้

*** =instructions
กำหนดให้ตัวแปร `theory` มีค่าเป็น `"factors for categorical variables"`.

*** =hint
แค่ใช้คำสั่งกำหนดค่า (`<-`) และตรวจสอบให้แน่ใจว่าคุณได้พิมพ์ตัวพิมพ์เล็กและตัวพิมพ์ใหญ่อย่างถูกต้องแล้ว

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# Assign to the variable theory what this chapter is about!
```

*** =solution
```{r}
# Assign to the variable theory what this chapter is about!
theory <- "factors for categorical variables"
```

*** =sct
```{r}
# MC-note: ideally, we could check for commonly mistyped variable names
#test_or({
#    bad_names <- c('Theory', "teory", "thoery", "theroy", "theiory", 'gender', 'value')
#    lapply(bad_names, test_object, eval=FALSE)
#    })

msg_undef <- paste("&#3604;&#3641;&#3648;&#3627;&#3617;&#3639;&#3629;&#3609;&#3623;&#3656;&#3634;&#3588;&#3640;&#3603;&#3618;&#3633;&#3591;&#3652;&#3617;&#3656;&#3652;&#3604;&#3657;&#3626;&#3619;&#3657;&#3634;&#3591;&#3605;&#3633;&#3623;&#3649;&#3611;&#3619;", "`theory`")
# ดูเหมือนว่าคุณยังไม่ได้สร้างตัวแปร `theory`

msg_incor <- paste("&#3605;&#3619;&#3623;&#3592;&#3626;&#3629;&#3610;&#3651;&#3627;&#3657;&#3649;&#3609;&#3656;&#3651;&#3592;&#3623;&#3656;&#3634;&#3588;&#3640;&#3603;&#3652;&#3604;&#3657;&#3585;&#3635;&#3627;&#3609;&#3604;&#3588;&#3656;&#3634;&#3586;&#3629;&#3591;&#3605;&#3633;&#3623;&#3649;&#3611;&#3619;", "`theory`", "&#3651;&#3627;&#3657;&#3648;&#3607;&#3656;&#3634;&#3585;&#3633;&#3610;", "`\"factors for categorical variables\"`", "&#40;&#3605;&#3657;&#3629;&#3591;&#3648;&#3611;&#3655;&#3609;&#3605;&#3633;&#3623;&#3614;&#3636;&#3617;&#3614;&#3660;&#3648;&#3621;&#3655;&#3585;&#3607;&#3633;&#3657;&#3591;&#3627;&#3617;&#3604;&#41;")
# ตรวจสอบให้แน่ใจว่าคุณได้กำหนดค่าของตัวแปร `theory` ให้เท่ากับ `\"factors for categorical variables\"` (ต้องเป็นตัวพิมพ์เล็กทั้งหมด)

msg_err <- paste("&#3605;&#3619;&#3623;&#3592;&#3626;&#3629;&#3610;&#3651;&#3627;&#3657;&#3649;&#3609;&#3656;&#3651;&#3592;&#3623;&#3656;&#3634;&#3588;&#3640;&#3603;&#3652;&#3604;&#3657;&#3626;&#3619;&#3657;&#3634;&#3591;&#3605;&#3633;&#3623;&#3649;&#3611;&#3619;", "`theory", "&#3629;&#3618;&#3656;&#3634;&#3591;&#3606;&#3641;&#3585;&#3605;&#3657;&#3629;&#3591;&#3650;&#3604;&#3618;&#3651;&#3594;&#3657;&#3588;&#3635;&#3626;&#3633;&#3656;&#3591;", "`<-`", "&#3651;&#3609;&#3585;&#3634;&#3619;&#3585;&#3635;&#3627;&#3609;&#3604;&#3588;&#3656;&#3634;&#3605;&#3633;&#3623;&#3649;&#3611;&#3619;")
# ตรวจสอบให้แน่ใจว่าคุณได้สร้างตัวแปร `theory` อย่างถูกต้องโดยใช้คำสั่ง `<-` ในการกำหนดค่าตัวแปร

success_msg <- "&#3648;&#3618;&#3637;&#3656;&#3618;&#3617;&#3617;&#3634;&#3585;&#33;&#32;&#3614;&#3619;&#3657;&#3629;&#3617;&#3607;&#3637;&#3656;&#3592;&#3632;&#3648;&#3619;&#3637;&#3618;&#3609;&#3619;&#3641;&#3657;&#3627;&#3619;&#3639;&#3629;&#3618;&#3633;&#3591;&#63;&#32;&#3648;&#3619;&#3634;&#3652;&#3611;&#3604;&#3641;&#3649;&#3610;&#3610;&#3613;&#3638;&#3585;&#3627;&#3633;&#3604;&#3605;&#3656;&#3629;&#3652;&#3611;&#3585;&#3633;&#3609;&#3648;&#3621;&#3618;"
# เยี่ยมมาก! พร้อมที่จะเรียนรู้หรือยัง? เราไปดูแบบฝึกหัดต่อไปกันเลย

# If get error and theory is undefined, point out the error
test_or(test_error(msg_err), test_object("theory", eval = FALSE))

test_object("theory", undefined_msg = msg_undef, incorrect_msg = msg_incor)
success_msg(success_msg)
```


--- type:NormalExercise xp:100 skills:1 key:6cc21c842b075347926bb1b244782213df32e370
## factor คืออะไรและทำไมเราถึงต้องใช้มัน? (2)

คุณสามารถใช้ function [`factor()`](http://www.rdocumentation.org/packages/base/functions/factor) ในการสร้าง factor ใน R ได้ สิ่งแรกที่คุณต้องทำก็คือสร้าง vector ของตัวแปรที่ใช้ในการแบ่งกลุ่ม ยกตัวอย่างเช่น `gender_vector` ที่เก็บข้อมูลเพศของคน 5 คนเอาไว้ดังตัวอย่าง:

```
gender_vector <- c("Male","Female","Female","Male","Male")
```

เราสามารถเห็นได้อย่างชัดเจนว่าข้อมูลถูกแบ่งออกเป็น 2 กลุ่มด้วยกัน (หรือในภาษา R เรียกว่า **'factor levels'**) ได้แก่ "Male" และ "Female".

function [`factor()`](http://www.rdocumentation.org/packages/base/functions/factor) จะทำการเปลี่ยนข้อมูลใน vector ให้เป็น factor:

```
factor_gender_vector <- factor(gender_vector)
```

*** =instructions
- แปลง vector ของข้อมูลแบบตัวอักษร `gender_vector` ให้เป็นข้อมูลแบบ factor ด้วย function `factor()` จากนั้นเก็บผลลัพธ์ไว้ในตัวแปร `factor_gender_vector`
- แสดงค่า `factor_gender_vector` และลองตรวจสอบดูว่า R ได้แสดง factor levels ไว้ใต้ค่าที่ถูกแสดงออกมา

*** =hint
แค่ใช้ function [`factor()`](http://www.rdocumentation.org/packages/base/functions/factor) กับ `gender_vector` เท่านั้น คุณก็จะได้ส่วนหนึ่งของคำตอบ ลองดูที่แบบฝึกหัดให้ดีๆอีกครั้ง คำตอบอยู่ในนั้นแล้ว...

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# Gender vector
gender_vector <- c("Male", "Female", "Female", "Male", "Male")

# Convert gender_vector to a factor
factor_gender_vector <-

# Print out factor_gender_vector

```

*** =solution
```{r}
# Gender vector
gender_vector <- c("Male", "Female", "Female", "Male", "Male")

# Convert gender_vector to a factor
factor_gender_vector <- factor(gender_vector)

# Print out factor_gender_vector
factor_gender_vector
```

*** =sct
```{r}
object_msg <- paste("&#3588;&#3640;&#3603;&#3652;&#3604;&#3657;&#3585;&#3635;&#3627;&#3609;&#3604;&#3651;&#3627;&#3657;&#3605;&#3633;&#3623;&#3649;&#3611;&#3619;", "`factor_gender_vector`", "&#3617;&#3637;&#3588;&#3656;&#3634;&#3648;&#3607;&#3656;&#3634;&#3585;&#3633;&#3610;&#3586;&#3657;&#3629;&#3617;&#3641;&#3621;&#3649;&#3610;&#3610;&#32;&#102;&#97;&#99;&#116;&#111;&#114;&#32;&#3592;&#3634;&#3585;", "`gender_vector`", "&#3627;&#3619;&#3639;&#3629;&#3618;&#3633;&#3591;&#63;")
# คุณได้กำหนดให้ตัวแปร `factor_gender_vector` มีค่าเท่ากับข้อมูลแบบ factor จาก `gender_vector` หรือยัง?

output_msg <- paste("&#3629;&#3618;&#3656;&#3634;&#3621;&#3639;&#3617;&#3626;&#3633;&#3656;&#3591;&#3649;&#3626;&#3604;&#3591;&#3588;&#3656;&#3634;", "`factor_gender_vector`!")
# อย่าลืมสั่งแสดงค่า `factor_gender_vector`!

success_msg <- paste("&#3648;&#3618;&#3637;&#3656;&#3618;&#3617;&#33;&#32;&#3606;&#3657;&#3634;&#3588;&#3640;&#3603;&#3605;&#3657;&#3629;&#3591;&#3585;&#3634;&#3619;&#3588;&#3635;&#3629;&#3608;&#3636;&#3610;&#3634;&#3618;&#3648;&#3614;&#3636;&#3656;&#3617;&#3648;&#3605;&#3636;&#3617;&#3648;&#3585;&#3637;&#3656;&#3618;&#3623;&#3585;&#3633;&#3610;&#32;&#102;&#117;&#110;&#99;&#116;&#105;&#111;&#110;", "`factor()`", "&#3649;&#3588;&#3656;&#3588;&#3640;&#3603;&#3614;&#3636;&#3617;&#3614;&#3660;", "`?factor`", "&#3621;&#3591;&#3652;&#3611;&#3651;&#3609;&#32;&#67;&#111;&#110;&#115;&#111;&#108;&#101;&#32;&#3585;&#3655;&#3592;&#3632;&#3648;&#3611;&#3655;&#3609;&#3585;&#3634;&#3619;&#3648;&#3611;&#3636;&#3604;&#3627;&#3609;&#3657;&#3634;&#32;&#104;&#101;&#108;&#112;&#32;&#3595;&#3638;&#3656;&#3591;&#3592;&#3632;&#3617;&#3637;&#3588;&#3635;&#3629;&#3608;&#3636;&#3610;&#3634;&#3618;&#3648;&#3614;&#3636;&#3656;&#3617;&#3648;&#3605;&#3636;&#3617;&#3629;&#3618;&#3641;&#3656;&#3651;&#3609;&#3609;&#3633;&#3657;&#3609;&#32;&#3649;&#3605;&#3656;&#3605;&#3629;&#3609;&#3609;&#3637;&#3657;&#32;&#3652;&#3611;&#3604;&#3641;&#3649;&#3610;&#3610;&#3613;&#3638;&#3585;&#3627;&#3633;&#3604;&#3605;&#3656;&#3629;&#3652;&#3611;&#3585;&#3633;&#3609;&#3604;&#3637;&#3585;&#3623;&#3656;&#3634;")
# เยี่ยม! ถ้าคุณต้องการคำอธิบายเพิ่มเติมเกี่ยวกับ function `factor()` แค่คุณพิมพ์ `?factor` ลงไปใน Console ก็จะเป็นการเปิดหน้า help ซึ่งจะมีคำอธิบายเพิ่มเติมอยู่ในนั้น แต่ตอนนี้ ไปดูแบบฝึกหัดต่อไปกันดีกว่า

test_object("factor_gender_vector", incorrect_msg = object_msg)
test_output_contains("factor_gender_vector", incorrect_msg = output_msg)
success_msg(success_msg);
```


--- type:NormalExercise xp:100 skills:1 key:5bd4f50afc2c2dbc881e16b8ca94ca56960dff42
## What's a factor and why would you use it? (3)

There are two types of categorical variables: a **nominal categorical variable** and an **ordinal categorical variable**.

A nominal variable is a categorical variable without an implied order. This means that it is impossible to say that 'one is worth more than the other'. For example, think of the categorical variable `animals_vector` with the categories `"Elephant"`, `"Giraffe"`, `"Donkey"` and `"Horse"`. Here, it is impossible to say that one stands above or below the other. (Note that some of you might disagree ;-) ).

In contrast, ordinal variables do have a natural ordering. Consider for example the categorical variable `temperature_vector` with the categories: `"Low"`, `"Medium"` and `"High"`. Here it is obvious that `"Medium"` stands above `"Low"`, and `"High"` stands above `"Medium"`.

*** =instructions
Click 'Submit Answer' to check how R constructs and prints nominal and ordinal variables. Do not worry if you do not understand all the code just yet, we will get to that.

*** =hint
Just click the 'Submit Answer' button and look at the console. Notice how R indicates the ordering of the factor levels for ordinal categorical variables.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# Animals
animals_vector <- c("Elephant", "Giraffe", "Donkey", "Horse")
factor_animals_vector <- factor(animals_vector)
factor_animals_vector

# Temperature
temperature_vector <- c("High", "Low", "High","Low", "Medium")
factor_temperature_vector <- factor(temperature_vector, order = TRUE, levels = c("Low", "Medium", "High"))
factor_temperature_vector
```

*** =solution
```{r}
# Animals
animals_vector <- c("Elephant", "Giraffe", "Donkey", "Horse")
factor_animals_vector <- factor(animals_vector)
factor_animals_vector

# Temperature
temperature_vector <- c("High", "Low", "High","Low", "Medium")
factor_temperature_vector <- factor(temperature_vector, order = TRUE, levels = c("Low", "Medium", "High"))
factor_temperature_vector
```

*** =sct
```{r}
msg <- "Do not change anything about the sample code. Simply hit the Submit Answer button and inspect the solution!"
test_object("animals_vector", undefined_msg = msg, incorrect_msg = msg)
test_object("temperature_vector", undefined_msg = msg, incorrect_msg = msg)
test_object("factor_animals_vector", undefined_msg = msg, incorrect_msg = msg)
test_output_contains("factor_animals_vector", incorrect_msg = msg)
test_object("factor_temperature_vector", undefined_msg = msg, incorrect_msg = msg)
test_output_contains("factor_temperature_vector", incorrect_msg = msg)
success_msg("Can you already tell what's happening in this exercise? Awesome! Continue to the next exercise and get into the details of factor levels.")
```


--- type:NormalExercise xp:100 skills:1 key:1aa698978d32d1a0befa4700d7da85a648e1d69e
## Factor levels

When you first get a data set, you will often notice that it contains factors with specific factor levels. However, sometimes you will want to change the names of these levels for clarity or other reasons. R allows you to do this with the function [`levels()`](http://www.rdocumentation.org/packages/base/functions/levels):

```
levels(factor_vector) <- c("name1", "name2",...)
```

A good illustration is the raw data that is provided to you by a survey. A standard question for every questionnaire is the gender of the respondent. You remember from the previous question that this is a factor and when performing the questionnaire on the streets its levels are often coded as `"M"` and `"F"`.

```
survey_vector <- c("M", "F", "F", "M", "M")
```

Next, when you want to start your data analysis, your main concern is to keep a nice overview of all the variables and what they mean. At that point, you will often want to change the factor levels to `"Male"` and `"Female"` instead of `"M"` and `"F"` to make your life easier.

**Watch out:** the order with which you assign the levels is important. If you type `levels(factor_survey_vector)`, you'll see that it outputs `[1] "F" "M"`. If you don't specify the levels of the factor when creating the vector, `R` will automatically assign them alphabetically. To correctly map `"F"` to `"Female"` and `"M"` to `"Male"`, the levels should be set to `c("Female", "Male")`, in this order order.

*** =instructions
- Check out the code that builds a factor vector from `survey_vector`. You should use `factor_survey_vector` in the next instruction.
- Change the factor levels of `factor_survey_vector` to `c("Female", "Male")`. Mind the order of the vector elements here.

*** =hint
Mind the order in which you have to type in the factor levels. Hint: look at the order in which the levels are printed when typing `levels(factor_survey_vector)`.

*** =pre_exercise_code
```{r}
# no pec
survey_vector <- c("M", "F", "F", "M", "M")
factor_survey_vector <- factor(survey_vector)
```

*** =sample_code
```{r}
# Code to build factor_survey_vector
survey_vector <- c("M", "F", "F", "M", "M")
factor_survey_vector <- factor(survey_vector)

# Specify the levels of factor_survey_vector
levels(factor_survey_vector) <-

factor_survey_vector
```

*** =solution
```{r}
# Code to build factor_survey_vector
survey_vector <- c("M", "F", "F", "M", "M")
factor_survey_vector <- factor(survey_vector)

# Specify the levels of factor_survey_vector
levels(factor_survey_vector) <- c("Female", "Male")

factor_survey_vector
```

*** =sct
```{r}
msg = "Do not change the definition of `survey_vector`!"
test_object("survey_vector", undefined_msg = msg, incorrect_msg = msg)
msg = "Do not change or remove the code to create the factor vector."
test_function("factor", "x", not_called_msg = msg, incorrect_msg = msg)

# MC-note: ideally would want to test assign operator `<-`, and have it highlight whole line.

# MC-note: or negate this test_student_typed, to highlight where they type this incorrect phrase
# test_student_typed('c("Male", "Female")')

test_object("factor_survey_vector", eq_condition = "equal",
            incorrect_msg = paste("Did you assign the correct factor levels to `factor_survey_vector`? Use `levels(factor_survey_vector) <- c(\"Female\", \"Male\")`. Remember that R is case sensitive!"))


success_msg("Wonderful! Proceed to the next exercise.")
```


--- type:NormalExercise xp:100 skills:1 key:a549f13c0644ccc89cd39a10aa48706754637ed0
## Summarizing a factor

After finishing this course, one of your favorite functions in R will be [`summary()`](http://www.rdocumentation.org/packages/base/functions/summary). This will give you a quick overview of the contents of a variable:

```
summary(my_var)
```

Going back to our survey, you would like to know how many `"Male"` responses you have in your study, and how many `"Female"` responses. The [`summary()`](http://www.rdocumentation.org/packages/base/functions/summary) function gives you the answer to this question.

*** =instructions
Ask a [`summary()`](http://www.rdocumentation.org/packages/base/functions/summary) of the `survey_vector` and `factor_survey_vector`. Interpret the results of both vectors. Are they both equally useful in this case?

*** =hint
Call the [`summary()`](http://www.rdocumentation.org/packages/base/functions/summary) function on both `survey_vector` and `factor_survey_vector`, it's as simple as that!

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# Build factor_survey_vector with clean levels
survey_vector <- c("M", "F", "F", "M", "M")
factor_survey_vector <- factor(survey_vector)
levels(factor_survey_vector) <- c("Female", "Male")
factor_survey_vector

# Generate summary for survey_vector


# Generate summary for factor_survey_vector

```

*** =solution
```{r}
# Build factor_survey_vector with clean levels
survey_vector <- c("M", "F", "F", "M", "M")
factor_survey_vector <- factor(survey_vector)
levels(factor_survey_vector) <- c("Female", "Male")
factor_survey_vector

# Generate summary for survey_vector
summary(survey_vector)

# Generate summary for factor_survey_vector
summary(factor_survey_vector)
```

*** =sct
```{r}
msg = "Do not change anything about the first few lines that define `survey_vector` and `factor_survey_vector`."
test_object("survey_vector", undefined_msg = msg, incorrect_msg = msg)
test_object("factor_survey_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)
msg <- "Have you correctly used `summary()` to generate a summary for `%s`?"
test_output_contains("summary(survey_vector)", incorrect_msg = sprintf(msg, "survey_vector"))
test_output_contains("summary(factor_survey_vector)", incorrect_msg = sprintf(msg, "factor_survey_vector"))
success_msg("Nice! Have a look at the output. The fact that you identified `\"Male\"` and `\"Female\"` as factor levels in `factor_survey_vector` enables R to show the number of elements for each category.")
```


--- type:NormalExercise xp:100 skills:1 key:90ecc160d1ebf2f75bf53f9c3843fc1632bdd0a5
## Battle of the sexes

In `factor_survey_vector` we have a factor with two levels: Male and Female. But how does R value these relatively to each other? In other words, who does R think is better, males or females?

*** =instructions
Read the code in the editor and click 'Submit Answer' to see whether males are worth more than females.

*** =hint
Just click 'Submit Answer' and have a look at output that gets printed to the console.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# Build factor_survey_vector with clean levels
survey_vector <- c("M", "F", "F", "M", "M")
factor_survey_vector <- factor(survey_vector)
levels(factor_survey_vector) <- c("Female", "Male")

# Male
male <- factor_survey_vector[1]

# Female
female <- factor_survey_vector[2]

# Battle of the sexes: Male 'larger' than female?
male > female
```

*** =solution
```{r}
# Build factor_survey_vector with clean levels
survey_vector <- c("M", "F", "F", "M", "M")
factor_survey_vector <- factor(survey_vector)
levels(factor_survey_vector) <- c("Female", "Male")

# Male
male <- factor_survey_vector[1]

# Female
female <- factor_survey_vector[2]

# Battle of the sexes: Male 'larger' than female?
male > female
```

*** =sct
```{r}
msg = "Do not change anything about the code; simply hit Submit Answer and look at the result."
test_object("survey_vector", undefined_msg = msg, incorrect_msg = msg)
test_object("factor_survey_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)
test_object("male", undefined_msg = msg, incorrect_msg = msg)
test_object("female", undefined_msg = msg, incorrect_msg = msg)
test_output_contains("male > female", incorrect_msg = msg)
success_msg("Phew, it seems that R is gender-neutral. Or maybe it just wants to stay out of these discussions ;-).")
```


--- type:NormalExercise xp:100 skills:1 key:9ab0928916bf84ab225713a9a1ce40d9e322c6a0
## Ordered factors

Since `"Male"` and `"Female"` are unordered (or nominal) factor levels, R returns a warning message, telling you that the greater than operator is not meaningful. As seen before, R attaches an equal value to the levels for such factors.

But this is not always the case! Sometimes you will also deal with factors that do have a natural ordering between its categories. If this is the case, we have to make sure that we pass this information to R...

Let us say that you are leading a research team of five data analysts and that you want to evaluate their performance. To do this, you track their speed, evaluate each analyst as `"slow"`, `"fast"` or `"insane"`, and save the results in `speed_vector`.

*** =instructions
As a first step, assign `speed_vector` a vector with 5 entries, one for each analyst. Each entry should be either `"slow"`, `"fast"`, or `"insane"`. Use the list below:

- Analyst 1 is fast,
- Analyst 2 is slow,
- Analyst 3 is slow,
- Analyst 4 is fast and
- Analyst 5 is insane.

No need to specify these are factors yet.

*** =hint
Assign to `speed_vector` a vector containing character strings, `"fast"`, `"slow"` ...

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# Create speed_vector
speed_vector <-
```

*** =solution
```{r}
# Create speed_vector
speed_vector <- c("fast", "slow", "slow", "fast", "insane")
```

*** =sct
```{r}
test_object("speed_vector",
            incorrect_msg = "`speed_vector` should be a vector with 5 entries, one for each analyst's speed rating. Don't use capital letters; R is case sensitive!")
success_msg("A job well done! Continue to the next exercise.")
```


--- type:NormalExercise xp:100 skills:1 key:279077d10248ce03d5f972939ef8576430a16683
## Ordered factors (2)

`speed_vector` should be converted to an ordinal factor since its categories have a natural ordering. By default, the function [`factor()`](http://www.rdocumentation.org/packages/base/functions/factor) transforms `speed_vector` into an unordered factor. To create an ordered factor, you have to add two additional arguments: `ordered` and `levels`.

```
factor(some_vector,
       ordered = TRUE,
       levels = c("lev1", "lev2" ...))
```

By setting the argument `ordered` to `TRUE` in the function [`factor()`](http://www.rdocumentation.org/packages/base/functions/factor), you indicate that the factor is ordered. With the argument `levels` you give the values of the factor in the correct order.

*** =instructions
From `speed_vector`, create an ordered factor vector: `factor_speed_vector`. Set `ordered` to `TRUE`, and set `levels` to `c("slow", "fast", "insane")`.

*** =hint
Use the function [`factor()`](http://www.rdocumentation.org/packages/base/functions/factor) to create `factor_speed_vector` based on `speed_character_vector`. The argument `ordered` should be set to `TRUE` since there is a natural ordering. Also, set `levels = c("slow", "fast", "insane")`.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# Create speed_vector
speed_vector <- c("fast", "slow", "slow", "fast", "insane")

# Convert speed_vector to ordered factor vector
factor_speed_vector <-

# Print factor_speed_vector
factor_speed_vector
summary(factor_speed_vector)
```

*** =solution
```{r}
# Create speed_vector
speed_vector <- c("fast", "slow", "slow", "fast", "insane")

# Add your code below
factor_speed_vector <- factor(speed_vector, ordered = TRUE, levels = c("slow", "fast", "insane"))

# Print factor_speed_vector
factor_speed_vector
summary(factor_speed_vector)
```

*** =sct
```{r}
msg = "Do not change anything about the command that specifies the `speed_vector` variable."
test_object("speed_vector", undefined_msg = msg, incorrect_msg = msg)
test_function("factor", args = c("x", "ordered", "levels"),
              incorrect_msg = c("The first argument you pass to `factor()` should be `speed_vector`.",
                                "Make sure to set `ordered = TRUE` inside your call of `factor()`.",
                                "Make sure to set `levels = c(\"slow\", \"fast\", \"insane\")` inside your call of `factor()`."))
test_object("factor_speed_vector", eq_condition = "equal",
            incorrect_msg = "There's still something wrong with `factor_speed_vector`; make sure to only pass `speed_vector`, `ordered = TRUE` and `levels = c(\"slow\", \"fast\", \"insane\")` inside your call of `factor()`.")
success_msg("Great! Have a look at the console. It is now indicated that the Levels indeed have an order associated, with the `<` sign. Continue to the next exercise.")
```


--- type:NormalExercise xp:100 skills:1 key:db16e69805625bcfde227743a8cbc985f8482a37
## Comparing ordered factors

Having a bad day at work, 'data analyst number two' enters your office and starts complaining that 'data analyst number five' is slowing down the entire project. Since you know that 'data analyst number two' has the reputation of being a smarty-pants, you first decide to check if his statement is true.

The fact that `factor_speed_vector` is now ordered enables us to compare different elements (the data analysts in this case). You can simply do this by using the well-known operators.

*** =instructions
- Use `[2]` to select from `factor_speed_vector` the factor value for the second data analyst. Store it as `da2`.
- Use `[5]` to select the `factor_speed_vector` factor value for the fifth data analyst. Store it as `da5`.
- Check if `da2` is greater than `da5`; simply print out the result. Remember that you can use the `>` operator to check whether one element is larger than the other.

*** =hint
- To select the factor value for the third data analyst, you'd need `factor_speed_vector[3]`.
- To compare two values, you can use `>`. For example: `da3 > da4`.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# Create factor_speed_vector
speed_vector <- c("fast", "slow", "slow", "fast", "insane")
factor_speed_vector <- factor(speed_vector, ordered = TRUE, levels = c("slow", "fast", "insane"))

# Factor value for second data analyst
da2 <-

# Factor value for fifth data analyst
da5 <-

# Is data analyst 2 faster than data analyst 5?

```

*** =solution
```{r}
# Create factor_speed_vector
speed_vector <- c("fast", "slow", "slow", "fast", "insane")
factor_speed_vector <- factor(speed_vector, ordered = TRUE, levels = c("slow", "fast", "insane"))

# Factor value for second data analyst
da2 <- factor_speed_vector[2]

# Factor value for fifth data analyst
da5 <- factor_speed_vector[5]

# Is data analyst 2 faster data analyst 5?
da2 > da5
```

*** =sct
```{r}
msg = "Do not change anything about the commands that define `speed_vector` and `factor_speed_vector`!"
test_object("speed_vector", undefined_msg = msg, incorrect_msg = msg)
test_object("factor_speed_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)

msg <- "Have you correctly selected the factor value for the %s data analyst? You can use `factor_speed_vector[%s]`."
test_object("da2", eq_condition = "equal", incorrect_msg = sprintf(msg, "second", "2"))
test_object("da5", eq_condition = "equal", incorrect_msg = sprintf(msg, "fifth", "5"))
test_output_contains("da2 > da5", incorrect_msg = "Have you correctly compared `da2` and `da5`? You can use the `>`. Simply print out the result.")

success_msg("Bellissimo! What do the result tell you? Data analyst two is complaining about the data analyst five while in fact he or she is the one slowing everything down! This concludes the chapter on factors. With a solid basis in vectors, matrices and factors, you're ready to dive into the wonderful world of data frames, a very important data structure in R!")
```



