--- 
title_meta  : บทที่ 2
title       : Vectors
description : "ในบทนี้ เราจะพาคุณไปยัง Las Vegas เพื่อเรียนรู้วิธีการวิเคราะห์ผลการแพ้ / ชนะพนันโดยใช้ R! เมื่อคุณเรียนรู้เนื้อหาทั้งหมดแล้ว คุณจะสามารถสร้าง vector และเปรียบเทียบ vector ที่แตกต่างกันได้!"

--- type:NormalExercise xp:100 skills:1 key:d9b453dbdd
## การสร้าง vector

Las Vegas มีอีกชื่อหนึ่งว่าเมืองแห่งบาป แต่บางคนก็เรียกมันว่า *สวรรค์ของนักสถิติ*!

ถ้าคุณมีความสามารถในการวิเคราะห์ข้อมูล (ซึ่งนั่นคือสิ่งที่ R ทำได้ดี) คุณอาจจะสามารถเป็นนักพนันที่ประสบความสำเร็จได้ ในบทเรียนนี้เราจะพาคุณไปดูวิธีการวิเคราะห์ผลแพ้ชนะการพนันที่เกิดขึ้นในอดีตซึ่งจะเป็นประโยชน์อย่างมากในการพัฒนาตัวเองในอนาคต ถ้าพร้อมแล้วก็ไปกันเลย มุ่งหน้าสู่ Vegas!

*** =instructions
- ยังจำสิ่งที่ได้เรียนไปในบทที่ 1 ได้ไหม? ให้คุณกำหนดให้ตัวแปร `vegas` มีค่าเป็น `"Go!"` อย่าลืมว่า R เป็นภาษาที่มีความเป็น case sensitive! (มองว่าตัวอักษรพิมพ์เล็กและพิมพ์ใหญ่มีค่าไม่เท่ากัน)

*** =hint
แค่พิมพ์โค้ดนี้ลงไปใน editor:
```
vegas <- "Go!"
```

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# Define the variable vegas
vegas <- 
```

*** =solution
```{r}
# Define the variable vegas
vegas <- "Go!"
```

*** =sct
```{r}
incorrect_msg <- paste("&#3605;&#3619;&#3623;&#3592;&#3626;&#3629;&#3610;&#3651;&#3627;&#3657;&#3649;&#3609;&#3656;&#3651;&#3592;&#3623;&#3656;&#3634;&#3588;&#3640;&#3603;&#3652;&#3604;&#3657;&#3585;&#3635;&#3627;&#3609;&#3604;&#3588;&#3656;&#3634;&#3607;&#3637;&#3656;&#3606;&#3641;&#3585;&#3605;&#3657;&#3629;&#3591;&#3651;&#3627;&#3657;&#3585;&#3633;&#3610;&#3605;&#3633;&#3623;&#3649;&#3611;&#3619;", "`vegas`", "&#40;&#3605;&#3633;&#3623;&#3614;&#3636;&#3617;&#3614;&#3660;&#3648;&#3621;&#3655;&#3585;&#3607;&#3633;&#3657;&#3591;&#3627;&#3617;&#3604;&#41;")
# ตรวจสอบให้แน่ใจว่าคุณได้กำหนดค่าที่ถูกต้องให้กับตัวแปร `vegas` (ตัวพิมพ์เล็กทั้งหมด)

success_msg <- "&#3648;&#3618;&#3637;&#3656;&#3618;&#3617;&#3617;&#3634;&#3585;&#33;&#32;&#3612;&#3656;&#3634;&#3609;&#3652;&#3611;&#3649;&#3610;&#3610;&#3613;&#3638;&#3585;&#3627;&#3633;&#3604;&#3605;&#3656;&#3629;&#3652;&#3611;&#3652;&#3604;&#3657;&#3648;&#3621;&#3618;"
# เยี่ยมมาก! ผ่านไปแบบฝึกหัดต่อไปได้เลย

test_object("vegas", incorrect_msg = incorrect_msg)
success_msg(success_msg)
```


--- type:NormalExercise xp:100 skills:1 key:fd427db76f
## การสร้าง vector (2)

ลองตั้งใจดูให้ดี!

vector ก็คือการนำข้อมูลมาร้อยเรียงกันเป็นแถวซึ่งสามารถเก็บข้อมูลได้ทุกประเภท ไม่ว่าจะเป็น ข้อมูลตัวเลข(numeric) ข้อมูลตัวอักษร(character) หรือข้อมูลแบบตรรกะ(logical) คุณจะได้ใช้ vector บ่อยมากในการวิเคราะห์ข้อมูลด้วย R ในตัวอย่างนี้เราจะให้คุณลองเก็บข้อมูลปะวัติการแพ้ / ชนะของคุณในการพนันที่คาสิโนลงไปใน vector

คุณสามารถสั่งให้ R สร้าง vector ได้โดยใช้คำสั่งรวมข้อมูลหรือ [`c()`](http://www.rdocumentation.org/packages/base/functions/c) (สามารถคลิกเพื่ออ่านคำอธิบายเพิ่มเติมได้) ขอเพียงคุณใส่ข้อมูลที่ต้องการจะนำมาสร้าง vector ลงไปในวงเล็บ แล้วคั่นแต่ละตัวด้วยเครื่องหมาย comma (`,`) ยกตัวอย่างเช่น:

```
numeric_vector <- c(1, 2, 3)
character_vector <- c("a", "b", "c")
```

และเมื่อคุณสร้าง vector ได้สำเร็จ คุณจะสามารถนำ vector เหล่านี้ไปใช้ในการคำนวณต่อๆไปได้

*** =instructions 
พิมพ์โค้ดให้เสร็จเรียบร้อย โดยการเก็บข้อมูล 3 ตัวเหล่านี้: `TRUE`, `FALSE` และ `TRUE` (เรียงตามลำดับ) ไว้ในตัวแปรชื่อ `boolean_vector` 

*** =hint 
กำหนดให้ตัวแปร `boolean_vector` มีค่าเท่ากับ `c(TRUE, FALSE, TRUE)` (เราสามารถใช้เครื่องหมาย `<-` ในการเก็บ vector ไว้ในตัวแปรได้เหมือนข้อมูลอื่นๆตามปกติ)

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
numeric_vector <- c(1, 10, 49)
character_vector <- c("a", "b", "c")

# Complete the code for boolean_vector
boolean_vector <-
```

*** =solution
```{r}
numeric_vector <- c(1, 10, 49)
character_vector <- c("a", "b", "c")

# Complete the code for boolean_vector
boolean_vector <- c(TRUE, FALSE, TRUE)
```

*** =sct
```{r}
msg <- paste("&#3629;&#3618;&#3656;&#3634;&#3648;&#3611;&#3621;&#3637;&#3656;&#3618;&#3609;&#3650;&#3588;&#3657;&#3604;&#3607;&#3637;&#3656;&#3585;&#3635;&#3627;&#3609;&#3604;&#3651;&#3627;&#3657;&#3585;&#3633;&#3610;", "`numeric_vector`", "&#3649;&#3621;&#3632;", "`character_vector`", "!")
# อย่าเปลี่ยนโค้ดที่กำหนดให้กับ `numeric_vector` และ `character_vector`!

incorrect_msg <- paste("&#3605;&#3619;&#3623;&#3592;&#3626;&#3629;&#3610;&#3651;&#3627;&#3657;&#3649;&#3609;&#3656;&#3651;&#3592;&#3623;&#3656;&#3634;&#3588;&#3640;&#3603;&#3652;&#3604;&#3657;&#3585;&#3635;&#3627;&#3609;&#3604;&#3588;&#3656;&#3634;&#3607;&#3637;&#3656;&#3606;&#3641;&#3585;&#3605;&#3657;&#3629;&#3591;&#3651;&#3627;&#3657;&#3585;&#3633;&#3610;", "`boolean_vector`", "&#3588;&#3656;&#3634;&#3607;&#3637;&#3656;&#3606;&#3641;&#3585;&#3605;&#3657;&#3629;&#3591;&#3588;&#3623;&#3619;&#3648;&#3611;&#3655;&#3609;", "`c(TRUE, FALSE, TRUE)`", "&#3650;&#3604;&#3618;&#3652;&#3617;&#3656;&#3617;&#3637;&#3648;&#3588;&#3619;&#3639;&#3656;&#3629;&#3591;&#3627;&#3617;&#3634;&#3618;", "`\"`", "&#3651;&#3604;&#3654;&#3607;&#3633;&#3657;&#3591;&#3626;&#3636;&#3657;&#3609;&#33;&#32;&#3607;&#3637;&#3656;&#3626;&#3635;&#3588;&#3633;&#3597;&#3588;&#3639;&#3629;&#3585;&#3634;&#3619;&#3648;&#3619;&#3637;&#3618;&#3591;&#3621;&#3635;&#3604;&#3633;&#3610;&#3605;&#3657;&#3629;&#3591;&#3606;&#3641;&#3585;&#3605;&#3657;&#3629;&#3591;&#3604;&#3657;&#3623;&#3618;")
# ตรวจสอบให้แน่ใจว่าคุณได้กำหนดค่าที่ถูกต้องให้กับ `boolean_vector` ค่าที่ถูกต้องควรเป็น `c(TRUE, FALSE, TRUE)` โดยไม่มีเครื่องหมาย `\"` ใดๆทั้งสิ้น! ที่สำคัญคือการเรียงลำดับต้องถูกต้องด้วย

success_msg <- paste("&#3618;&#3629;&#3617;&#3648;&#3618;&#3637;&#3656;&#3618;&#3617;&#3626;&#3640;&#3604;&#3654;&#33;&#32;&#3585;&#3634;&#3619;&#3648;&#3614;&#3636;&#3656;&#3617;&#3623;&#3657;&#3609;&#3623;&#3619;&#3619;&#3588;&#3604;&#3657;&#3634;&#3609;&#3627;&#3621;&#3633;&#3591;&#32;&#99;&#111;&#109;&#109;&#97;&#32;&#3586;&#3657;&#3634;&#3591;&#3651;&#3609;&#32;&#102;&#117;&#110;&#99;&#116;&#105;&#111;&#110;", "`c()`", "&#3592;&#3632;&#3594;&#3656;&#3623;&#3618;&#3607;&#3635;&#3651;&#3627;&#3657;&#3650;&#3588;&#3657;&#3604;&#3586;&#3629;&#3591;&#3588;&#3640;&#3603;&#3629;&#3656;&#3634;&#3609;&#3591;&#3656;&#3634;&#3618;&#3586;&#3638;&#3657;&#3609;&#32;&#3648;&#3619;&#3634;&#3617;&#3634;&#3621;&#3629;&#3591;&#3613;&#3638;&#3585;&#3626;&#3619;&#3657;&#3634;&#3591;&#32;&#118;&#101;&#99;&#116;&#111;&#114;&#32;&#3585;&#3633;&#3609;&#3629;&#3637;&#3585;&#3626;&#3633;&#3585;&#3627;&#3609;&#3656;&#3629;&#3618;&#3604;&#3637;&#3652;&#3627;&#3617;&#63;")
# ยอมเยี่ยมสุดๆ! การเพิ่มว้นวรรคด้านหลัง comma ข้างใน function `c()` จะช่วยทำให้โค้ดของคุณอ่านง่ายขึ้น เรามาลองฝึกสร้าง vector กันอีกสักหน่อยดีไหม?

test_object("numeric_vector", undefined_msg = msg, incorrect_msg = msg)
test_object("character_vector", undefined_msg = msg, incorrect_msg = msg)
test_object("boolean_vector",
            incorrect_msg = "Make sure that you assign the correct values to `boolean_vector`. Use `c(TRUE, FALSE, TRUE)`. Don't place quotes around `TRUE` and `FALSE`! Also, make sure to adopt the same order as listed in the instructions.")
success_msg("Perfect! Notice that adding a space behind the commas in the `c()` function improves the readability of your code. Let's practice some more with vector creation in the next exercise.")
```


--- type:NormalExercise xp:100 skills:1 key:9f41229dbc
## การสร้าง vector (3)

1 สัปดาห์ผ่านไปกับการใช้ชีวิตใน Las Vegas แต่ดูเหมือนคุณจะทำเงินจากการพนันได้ไม่เท่าที่หวังไว้ในตอนแรก ดังนั้นคุณจึงตัดสินใจว่าถึงเวลาแล้วที่จะต้องหวังพึ่งพลังในการวิเคราะห์ข้อมูลด้วย R!

แต่ก่อนที่จะเริ่มทำการวิเคราะห์ คุณคิดว่าคุณอยากดูข้อมูลการแพ้ / ชนะพนัน (เป็นจำนวนเงิน) ทั้งหมดภายใน 1 สัปดาห์ที่ผ่านมาเสียก่อน:

สำหรับ `poker_vector`: 

- วันจันทร์ คุณได้เงิน $140
- วันอังคาร คุณเสียเงิน $50
- วันพุธ คุณได้เงิน $20 
- วันพฤหัส คุณเสียเงิน $120
- วันศุกร์ คุณได้เงิน $240

สำหรับ `roulette_vector`: 

- วันจันทร์ คุณได้เงิน $24
- วันอังคาร คุณเสียเงิน $50
- วันพุธ คุณได้เงิน $100
- วันพฤหัส คุณเสียเงิน $350
- วันศุกร์ คุณได้เงิน $10

เกมที่คุณล่นมีแค่ poker และ roulette ในการที่จะนำข้อมูลเหล่านี้ไปใช้ใน R คุณจะต้องสร้างตัวแปร `poker_vector` และ `roulette_vector`.

*** =instructions
เก็บ vector ของผลการแพ้ / ชนะพนันแต่ละครั้งในการเล่นเกม roulette ลงไปในตัวแปร `roulette_vector`.

*** =hint
จากตัวอย่างที่ให้ไว้ เรามีผลแพ้ / ชนะจากเกม poker เก็บไว้ในตัวแปร `poker_vector` คุณก็แค่ต้องนำตัวอย่างนี้ไปปรับเปลี่ยนให้เหมาะสมกับผลแพ้ / ชนะของเกม roulette และตัวแปร `roulette_vector` และอย่าลืมว่าเราให้เครื่องหมายลบกับการแพ้พนันเพราะเราเสียเงินในการเล่นครั้งนั้นๆ

*** =sample_code
```{r}
# Poker winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)

# Roulette winnings from Monday to Friday
roulette_vector <-  
```

*** =solution
```{r}
# Poker winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)

# Roulette winnings from Monday to Friday
roulette_vector <- c(-24, -50, 100, -350, 10)
```

*** =sct
```{r}
poker_incorrect <- paste("&#3605;&#3619;&#3623;&#3592;&#3626;&#3629;&#3610;&#3651;&#3627;&#3657;&#3649;&#3609;&#3656;&#3651;&#3592;&#3623;&#3656;&#3634;&#3588;&#3640;&#3603;&#3652;&#3604;&#3657;&#3585;&#3635;&#3627;&#3609;&#3604;&#3588;&#3656;&#3634;&#3607;&#3637;&#3656;&#3606;&#3641;&#3585;&#3605;&#3657;&#3629;&#3591;&#3651;&#3627;&#3657;", "`poker_vector`")
# ตรวจสอบให้แน่ใจว่าคุณได้กำหนดค่าที่ถูกต้องให้ `poker_vector`

roulette_incorrect <- paste("&#3605;&#3619;&#3623;&#3592;&#3626;&#3629;&#3610;&#3651;&#3627;&#3657;&#3649;&#3609;&#3656;&#3651;&#3592;&#3623;&#3656;&#3634;&#3588;&#3640;&#3603;&#3652;&#3604;&#3657;&#3585;&#3635;&#3627;&#3609;&#3604;&#3588;&#3656;&#3634;&#3607;&#3637;&#3656;&#3606;&#3641;&#3585;&#3605;&#3657;&#3629;&#3591;&#3651;&#3627;&#3657;", "`roulette_vector`", "&#3649;&#3621;&#3632;&#3586;&#3657;&#3629;&#3617;&#3641;&#3621;&#3605;&#3657;&#3629;&#3591;&#3648;&#3619;&#3637;&#3618;&#3591;&#3621;&#3635;&#3604;&#3633;&#3610;&#3629;&#3618;&#3656;&#3634;&#3591;&#3606;&#3641;&#3585;&#3605;&#3657;&#3629;&#3591;&#3604;&#3657;&#3623;&#3618;&#3609;&#3632;")
# ตรวจสอบให้แน่ใจว่าคุณได้กำหนดค่าที่ถูกต้องให้ `roulette_vector` และข้อมูลต้องเรียงลำดับอย่างถูกต้องด้วยนะ

success_msg <- "&#3604;&#3637;&#3617;&#3634;&#3585;&#33;&#32;&#3651;&#3609;&#3585;&#3634;&#3619;&#3648;&#3594;&#3655;&#3588;&#3588;&#3656;&#3634;&#3586;&#3629;&#3591;&#3586;&#3657;&#3629;&#3617;&#3641;&#3621;&#3651;&#3609;&#32;&#118;&#101;&#99;&#116;&#111;&#114;&#32;&#3592;&#3635;&#3652;&#3623;&#3657;&#3623;&#3656;&#3634;&#3648;&#3619;&#3634;&#3626;&#3634;&#3617;&#3634;&#3619;&#3606;&#3614;&#3636;&#3617;&#3614;&#3660;&#3594;&#3639;&#3656;&#3629;&#3605;&#3633;&#3623;&#3649;&#3611;&#3619;&#32;&#118;&#101;&#99;&#116;&#111;&#114;&#32;&#3609;&#3633;&#3657;&#3609;&#3654;&#3621;&#3591;&#3652;&#3611;&#3651;&#3609;&#32;&#67;&#111;&#110;&#115;&#111;&#108;&#101;&#32;&#3649;&#3621;&#3657;&#3623;&#3585;&#3604;&#32;&#69;&#110;&#116;&#101;&#114;&#32;&#3652;&#3604;&#3657;&#3648;&#3621;&#3618;"
# ดีมาก! จำไว้ว่าเราสามารถพิมพ์ชื่อตัวแปร vector ลงไปใน Console แล้วกด Enter เพื่อเช็คค่าของข้อมูลข้างใน vector นั้นๆได้เลย

test_object("poker_vector", incorrect_msg = poker_incorrect)
test_object("roulette_vector", incorrect_msg = roulette_incorrect)
success_msg(success_msg)
```


--- type:NormalExercise xp:100 skills:1 key:3b0b80b192
## การตั้งชื่อให้ข้อมูลภายใน vector

สำหรับนักวิเคราะห์ข้อมูล การเข้าใจและมองเห็นภาพรวมของข้อมูลที่คุณกำลังใช้อยู่เป็นเรื่องสำคัญมาก คุณควรจะเข้าใจด้วยว่าข้อมูลแต่ละตัวภายใน vector นั้นแสดงถึงอะไร

ในแบบฝึกหัดก่อนหน้านี้ เราได้เรียนรู้การสร้าง vector ที่เก็บข้อมูลการแพ้ / ชนะ พนันของคุณเอาไว้ ข้อมูลแต่ละตัว(หรือที่เราเรียกว่า element) ใน vector เป็นตัวแทนของจำนวนเงินที่คุณได้หรือเสียในแต่ละวัน แต่เรายังไม่มีอะไรที่บอกได้ว่าข้อมูลแต่ละตัวเป็นข้อมูลของวันไหน มันน่าจะเป็นเรื่องดีถ้าเราระบุชื่อวันให้กับข้อมูลภายใน vector ด้วย เพราะนั่นจะช่วยให้เราเข้าใจข้อมูลได้ง่ายขึ้น!

คุณสามารถตั้งชื่อให้กับข้อมูลแต่ละตัวภายใน vector ได้โดยการใช้ function `names()` ลองดูโค้ดด้านล่างเป็นตัวอย่าง:

```
some_vector <- c("John Doe", "poker player")
names(some_vector) <- c("Name", "Profession")
```

โค้ดเริ่มจากการสร้างตัวแปร `some_vector` เพื่อให้เก็บค่าของ vector ที่มีข้อมูล 2 ตัว จากนั้นเราตั้งชื่อข้อมูลตัวแรกว่า `Name` และตั้งชื่อให้ตัวที่ 2 ว่า `Profession` ถ้าเราสั่งให้ R แสดงค่าของ vector นี้ออกมา Console จะทำการแสดงผลตามด้านล่าง:

```
          Name     Profession 
    "John Doe" "poker player" 
```

*** =instructions
- โค้ดทางด้านขวามือทำการตั้งชื่อให้กับข้อมูลแต่ละตัวใน `poker_vector` ด้วยชื่อวันในสัปดาห์ ให้คุณลองทำแบบเดียวกันกับ `roulette_vector` ดู

*** =hint
คุณสามารถใช้คำสั่ง `names(roulette_vector)` เพื่อตั้งชื่อให้กับข้อมูลแต่ละตัวใน `roulette_vector` ได้ ตรวจสอบให้แน่ใจว่าคุณได้ตั้งชื่อวันอย่างถูกต้อง และอย่าลืมว่า R นั้นมีความเป็น case sensitive!

*** =sample_code
```{r}
# Poker winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)

# Roulette winnings from Monday to Friday
roulette_vector <- c(-24, -50, 100, -350, 10)

# Assign days as names of poker_vector
names(poker_vector) <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")

# Assign days as names of roulette_vectors
```

*** =solution
```{r}
# Poker winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)

# Roulette winnings from Monday to Friday
roulette_vector <- c(-24, -50, 100, -350, 10)

# Assign days as names of poker_vector
names(poker_vector) <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")

# Assign days as names of roulette_vectors
names(roulette_vector) <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
```

*** =sct
```{r}
poker_incorrect <- paste("&#3629;&#3618;&#3656;&#3634;&#3648;&#3611;&#3621;&#3637;&#3656;&#3618;&#3609;&#3588;&#3656;&#3634;&#32;&#118;&#101;&#99;&#116;&#111;&#114;&#32;&#3607;&#3637;&#3656;&#3585;&#3635;&#3627;&#3609;&#3604;&#3651;&#3627;&#3657;&#3585;&#3633;&#3610;&#3605;&#3633;&#3623;&#3649;&#3611;&#3619;", "`poker_vector`")
# อย่าเปลี่ยนค่า vector ที่กำหนดให้กับตัวแปร `poker_vector`

roulette_incorrect <- paste("&#3629;&#3618;&#3656;&#3634;&#3648;&#3611;&#3621;&#3637;&#3656;&#3618;&#3609;&#3588;&#3656;&#3634;&#32;&#118;&#101;&#99;&#116;&#111;&#114;&#32;&#3607;&#3637;&#3656;&#3585;&#3635;&#3627;&#3609;&#3604;&#3651;&#3627;&#3657;&#3585;&#3633;&#3610;&#3605;&#3633;&#3623;&#3649;&#3611;&#3619;", "`roulette_vector`")
# อย่าเปลี่ยนค่า vector ที่กำหนดให้กับตัวแปร `roulette_vector`

poker_incorrect_eq <- paste("&#3607;&#3637;&#3656;&#3588;&#3640;&#3603;&#3605;&#3657;&#3629;&#3591;&#3607;&#3635;&#3588;&#3639;&#3629;&#3649;&#3588;&#3656;&#3605;&#3633;&#3657;&#3591;&#3594;&#3639;&#3656;&#3629;&#3651;&#3627;&#3657;&#3649;&#3605;&#3656;&#3621;&#3632;&#32;&#101;&#108;&#101;&#109;&#101;&#110;&#116;&#32;&#3586;&#3629;&#3591;", "`roulette vector`", "&#3648;&#3607;&#3656;&#3634;&#3609;&#3633;&#3657;&#3609;&#32;&#3629;&#3618;&#3656;&#3634;&#3648;&#3611;&#3621;&#3637;&#3656;&#3618;&#3609;&#3594;&#3639;&#3656;&#3629;&#3607;&#3637;&#3656;&#3648;&#3619;&#3634;&#3605;&#3633;&#3657;&#3591;&#3651;&#3627;&#3657;&#3585;&#3633;&#3610;&#3649;&#3605;&#3656;&#3621;&#3632;&#32;&#101;&#108;&#101;&#109;&#101;&#110;&#116;&#32;&#3586;&#3629;&#3591;", "`poker_vector`")
# ที่คุณต้องทำคือแค่ตั้งชื่อให้แต่ละ element ของ `roulette vector` &#3651;&#3627;&#3657;&#3588;&#3640;&#3603;&#3651;&#3594;&#3657;&#32;&#118;&#101;&#99;&#116;&#111;&#114;&#32;&#3648;&#3604;&#3637;&#3618;&#3623;&#3585;&#3633;&#3609;&#3585;&#3633;&#3610;&#3607;&#3637;&#3656;&#3648;&#3619;&#3634;&#3651;&#3594;&#3657;&#3605;&#3633;&#3657;&#3591;&#3594;&#3639;&#3656;&#3629;&#3651;&#3627;&#3657;&#3586;&#3657;&#3629;&#3617;&#3641;&#3621;&#3651;&#3609;ท่านั้น อย่าเปลี่ยนชื่อที่เราตั้งให้กับแต่ละ element ของ `poker_vector`

roulette_incorrect_eq <- paste("&#3651;&#3627;&#3657;&#3588;&#3640;&#3603;&#3651;&#3594;&#3657;&#32;&#118;&#101;&#99;&#116;&#111;&#114;&#32;&#3648;&#3604;&#3637;&#3618;&#3623;&#3585;&#3633;&#3609;&#3585;&#3633;&#3610;&#3607;&#3637;&#3656;&#3648;&#3619;&#3634;&#3651;&#3594;&#3657;&#3605;&#3633;&#3657;&#3591;&#3594;&#3639;&#3656;&#3629;&#3651;&#3627;&#3657;&#3586;&#3657;&#3629;&#3617;&#3641;&#3621;&#3651;&#3609;", "`poker_vector`", "&#3651;&#3609;&#3585;&#3634;&#3619;&#3605;&#3633;&#3657;&#3591;&#3594;&#3639;&#3656;&#3629;&#3651;&#3627;&#3657;&#3585;&#3633;&#3610;&#3586;&#3657;&#3629;&#3617;&#3641;&#3621;&#3651;&#3609;", "`roulette_vector`")
# ให้คุณใช้ vector เดียวกันกับที่เราใช้ตั้งชื่อให้ข้อมูลใน `poker_vector` ในการตั้งชื่อให้กับข้อมูลใน `roulette_vector`

test_object("poker_vector", incorrect_msg = poker_incorrect)
test_object("roulette_vector", incorrect_msg = roulette_incorrect)
test_object("poker_vector", eq_condition = "equal", incorrect_msg = poker_incorrect_eq)
test_object("roulette_vector", eq_condition = "equal", incorrect_msg = roulette_incorrect_eq)
success_msg("Well done! Continue to the next exercise.")
```


--- type:NormalExercise xp:100 skills:1 key:6858c65a4a
## การตั้งชื่อให้ข้อมูลภายใน vector (2)

ถ้าคุณอยากเป็นนักสถิติที่ดี บางครั้งคุณก็ต้องทำงานอย่างขี้เกียจบ้าง (แต่ใช้วิธีที่ฉลาดนะ)If you want to become a good statistician, you have to become lazy. (If you are already lazy, chances are high you are one of those exceptional, natural-born statistical talents.)

ก่อนหน้านี้ In the previous exercises you probably experienced that it is boring and frustrating to type and retype information such as the days of the week. However, when you look at it from a higher perspective, there is a more efficient way to do this, namely, to assign the days of the week vector to a **variable**! 

Just like you did with your poker and roulette returns, you can also create a variable that contains the days of the week. This way you can use and re-use it.

*** =instructions
- A variable `days_vector` that contains the days of the week has already been created for you.
- Use `days_vector` to set the names of `poker_vector` and `roulette_vector`.

*** =hint
You can use `names(poker_vector) <- days_vector` to set the names of the elements `poker_vector`. Do a similar thing for `roulette_vector`.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# Poker winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)

# Roulette winnings from Monday to Friday
roulette_vector <- c(-24, -50, 100, -350, 10)

# The variable days_vector
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
 
# Assign the names of the day to roulette_vector and poker_vector
names(poker_vector) <-   
names(roulette_vector) <-
```

*** =solution
```{r}
# Poker winnings from Monday to Friday
poker_vector <- c(140, -50, 20, -120, 240)

# Roulette winnings from Monday to Friday
roulette_vector <- c(-24, -50, 100, -350, 10)

# The variable days_vector
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")

# Assign the names of the day to roulette_vector and poker_vector
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector
```

*** =sct
```{r}
msg <- "Do not changes the predefined variables `poker_vector`, `roulette_vector` or `days_vector`."
test_object("poker_vector", undefined_msg = msg, incorrect_msg = msg)
test_object("roulette_vector", undefined_msg = msg, incorrect_msg = msg)
test_object("days_vector", undefined_msg = msg, incorrect_msg = msg)

test_object("poker_vector",
            eq_condition = "equal",
            incorrect_msg = "Make sure that you assign `days_vector` to the names of `poker_vector`.")
test_object("roulette_vector",
            eq_condition = "equal",
            incorrect_msg = "Make sure that you assign `days_vector` to the names of `roulette_vector`.")
success_msg("Nice one! A word of advice: try to avoid code duplication at all times. Continue to the next exercise and learn how to do arithmetic with vectors!")
```


--- type:NormalExercise xp:100 skills:1 key:da995f099f
## Calculating total winnings

Now that you have the poker and roulette winnings nicely as named vectors, you can start doing some data analytical magic. 

You want to find out the following type of information:

- How much has been your overall profit or loss per day of the week?
- Have you lost money over the week in total?
- Are you winning/losing money on poker or on roulette?

To get the answers, you have to do arithmetic calculations on vectors. 

It is important to know is that if you sum two vectors in R, it takes the element-wise sum. For example, the following three statements are completely equivalent:

```
c(1, 2, 3) + c(4, 5, 6)
c(1 + 4, 2 + 5, 3 + 6)
c(5, 7, 9)
```

You can also do the calculations with variables that represent vectors:

```
a <- c(1, 2, 3) 
b <- c(4, 5, 6)
c <- a + b
```

*** =instructions
- Take the sum of the variables `A_vector` and `B_vector` and it assign to `total_vector`.
- Inspect the result by printing out `total_vector`.

*** =hint
Use the `+` operator to sum `A_vector` and `B_vector`. Use `<-` to assign the result to `total_vector`.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
A_vector <- c(1, 2, 3)
B_vector <- c(4, 5, 6)

# Take the sum of A_vector and B_vector
total_vector <- 
  
# Print out total_vector

```

*** =solution
```{r}
A_vector <- c(1, 2, 3)
B_vector <- c(4, 5, 6)

# Take the sum of A_vector and B_vector
total_vector <- A_vector + B_vector

# Print out total_vector
total_vector
```

*** =sct
```{r}
msg <- "Do not change the contents of `A_vector` or `B_vector`!"
test_object("A_vector", undefined_msg = msg, incorrect_msg = msg)
test_object("B_vector", undefined_msg = msg, incorrect_msg = msg)
test_object("total_vector", incorrect_msg = "Make sure that `total_vector` contains the sum of `A_vector` and `B_vector`.")
test_output_contains("total_vector", incorrect_msg = "Don't forget to print out `total_vector`! Simply write `total_vector` on a new line.")
success_msg("Good job! Continue to the next exercise.")
```


--- type:NormalExercise xp:100 skills:1 key:2969d8ed65
## Calculating total winnings (2)

Now you understand how R does arithmetic with vectors, it is time to get those Ferraris in your garage! First, you need to understand what the overall profit or loss per day of the week was. The total daily profit is the sum of the profit/loss you realized on poker per day, and the profit/loss you realized on roulette per day. 

In R, this is just the sum of `roulette_vector` and `poker_vector`.

*** =instructions
Assign to the variable `total_daily` how much you won or lost on each day in total (poker and roulette combined).

*** =hint
Similar to the previous exercise, assign the sum of two vectors to a new variable, `total_daily`.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# Poker and roulette winnings from Monday to Friday:
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Assign to total_daily how much you won/lost on each day
total_daily <- 
```

*** =solution
```{r}
# Poker and roulette winnings from Monday to Friday:
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Assign to total_daily how much you won/lost on each day
total_daily <- poker_vector + roulette_vector
```

*** =sct
```{r}
msg = "Do not change anything about the definition and naming of `poker_vector` and `roulette_vector`."
test_object("days_vector", undefined_msg = msg, incorrect_msg = msg)
test_object("poker_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)
test_object("roulette_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)

test_object("total_daily", incorrect_msg = "Make sure that you assign the sum of `poker_vector` and `roulette_vector` to `total_daily`.")

success_msg("Great! Continue to the next exercise.")
```


--- type:NormalExercise xp:100 skills:1 key:e66a56b9f0
## Calculating total winnings (3)

Based on the previous analysis, it looks like you had a mix of good and bad days. This is not what your ego expected, and you wonder if there may be a very tiny chance you have lost money over the week in total? 

A function that helps you to answer this question is [`sum()`](http://www.rdocumentation.org/packages/base/functions/sum). It calculates the sum of all elements of a vector. For example, to calculate the total amount of money you have lost/won with poker you do: 

```
total_poker <- sum(poker_vector)
```

*** =instructions
- Calculate the total amount of money that you have won/lost with roulette and assign to the variable `total_roulette`.
- Now that you have the totals for roulette and poker, you can easily calculate `total_week` (which is the sum of all gains and losses of the week).
- Print out `total_week`.

*** =hint
Use the [`sum()`](http://www.rdocumentation.org/packages/base/functions/sum) function to get the total of the `roulette_vector`. `total_week` is then the sum of `roulette_vector` and `poker_vector`.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# Poker and roulette winnings from Monday to Friday:
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Total winnings with poker
total_poker <- sum(poker_vector)

# Total winnings with roulette
total_roulette <-  

# Total winnings overall
total_week <- 

# Print out total_week
  
```

*** =solution
```{r}
# Poker and roulette winnings from Monday to Friday:
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Total winnings with poker
total_poker <- sum(poker_vector)

# Total winnings with roulette
total_roulette <-  sum(roulette_vector)

# Total winnings overall
total_week <- total_roulette + total_poker

# Print out total_week
total_week
```

*** =sct
```{r}
msg = "Do not change anything about the definition and naming of `poker_vector` and `roulette_vector`."
test_object("days_vector", undefined_msg = msg, incorrect_msg = msg)
test_object("poker_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)
test_object("roulette_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)
test_object("total_poker", 
            incorrect_msg = "Make sure that you assign to `total_poker` the sum of the `poker_vector`.")
test_object("total_roulette",
            incorrect_msg = "Make sure that you assign to `total_roulette` the sum of the `roulette_vector`.")
test_object("total_week",
            incorrect_msg = "Make sure that you assign to `total_week` the sum of the other two total vectors: `total_roulette` and `total_poker`.")

test_output_contains("total_week", incorrect_msg = "Don't forget to write `total_week` on a new line to print out the variable.")
success_msg("Well done. This is pretty bad news...")
```


--- type:NormalExercise xp:100 skills:1 key:f532f5332d
## Comparing total winnings

Oops, it seems like you are losing money. Time to rethink and adapt your strategy! This will require some deeper analysis... 

After a short brainstorm in your hotel's jacuzzi, you realize that a possible explanation might be that your skills in roulette are not as well developed as your skills in poker. So maybe your total gains in poker are higher (or `>` ) than in roulette.

*** =instructions
- Calculate `total_poker` and `total_roulette` as in the previous exercise. Use the `sum()` function twice.
- Check if your total gains in poker are higher than for roulette by using a comparison. Simply print out the result of this comparison. What do you conclude, should you focus on roulette or on poker?

*** =hint
- You partly calculated the answer to this question in the previous exercise already!
- To check if 6 is larger than 5, you type `6 > 5`. This returns a logical value (`TRUE` or `FALSE`).

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# Poker and roulette winnings from Monday to Friday:
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Calculate total gains for poker and roulette
total_poker <-
total_roulette <-

# Check if you realized higher total gains in poker than in roulette 

```

*** =solution
```{r}
# Poker and roulette winnings from Monday to Friday:
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Calculate total gains for poker and roulette
total_poker <- sum(poker_vector)
total_roulette <- sum(roulette_vector)

# Check if you realized higher total gains in poker than in roulette
total_poker > total_roulette
```

*** =sct
```{r}
msg <- "Do not change anything about the definition and naming of `poker_vector` and `roulette_vector`."
test_object("days_vector", undefined_msg = msg, incorrect_msg = msg)
test_object("poker_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)
test_object("roulette_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)

test_object("total_poker", 
            incorrect_msg = "Make sure that you assign to `total_poker` the sum of the `poker_vector`. Use `sum()`.")
test_object("total_roulette",
            incorrect_msg = "Make sure that you assign to `total_roulette` the sum of the `roulette_vector`. Use `sum()`.")
test_output_contains("total_poker > total_roulette",
                     incorrect_msg = "Have you correctly carried out the comparison? To check if `total_poker` is greater than `total_roulette`, you can use `total_poker > total_roulette`.")
success_msg("Good job! Continue to the next exercise.")
```


--- type:NormalExercise xp:100 skills:1 key:8d78be44e9
## Vector selection: the good times

Your hunch seemed to be right. It appears that the poker game is more your cup of tea than roulette. 

Another possible route for investigation is your performance at the beginning of the working week compared to the end of it. You did have a couple of Margarita cocktails at the end of the week... 

To answer that question, you only want to focus on a selection of the `total_vector`. In other words, our goal is to select specific elements of the vector. To select elements of a vector (and later matrices, data frames, ...), you can use square brackets. Between the square brackets, you indicate what elements to select. For example, to select the first element of the vector, you type `poker_vector[1]`. To select the second element of the vector, you type `poker_vector[2]`, etc. Notice that the first element in a vector has index 1, not 0 as in many other programming languages.

*** =instructions
Assign the poker results of Wednesday to the variable `poker_wednesday`.

*** =hint
Wednesday is the third element of `poker_vector`, and can thus be selected with `poker_vector[3]`.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# Poker and roulette winnings from Monday to Friday:
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Define a new variable based on a selection
poker_wednesday <- 
```

*** =solution
```{r}
# Poker and roulette winnings from Monday to Friday:
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Define a new variable based on a selection
poker_wednesday <- poker_vector[3]
```

*** =sct
```{r}
msg = "Do not change anything about the definition and naming of `poker_vector` and `roulette_vector`."
test_object("days_vector", undefined_msg = msg, incorrect_msg = msg)
test_object("poker_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)
test_object("roulette_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)
test_object("poker_wednesday", 
            undefined_msg = "Please make sure to define a variable `poker_wednesday`.",
            incorrect_msg = "It looks like `poker_wednesday` does not contain the correct value of the `poker_vector`.")
success_msg("Great! R also makes it possible to select multiple elements from a vector at once. Learn how in the next exercise!")
```


--- type:NormalExercise xp:100 skills:1 key:1351521670
## Vector selection: the good times (2)

How about analyzing your midweek results? 

To select multiple elements from a vector, you can add square brackets at the end of it. You can indicate between the brackets what elements should be selected. For example: suppose you want to select the first and the fifth day of the week: use the vector `c(1, 5)` between the square brackets. For example, the code below selects the first and fifth element of `poker_vector`:

```
poker_vector[c(1, 5)]
```

*** =instructions
Assign the poker results of Tuesday, Wednesday and Thursday to the variable `poker_midweek`.

*** =hint
Use the vector `c(2, 3, 4)` between square brackets to select the correct elements of `poker_vector`.

*** =pre_exercise_code
```{r}
# no pec
``` 

*** =sample_code
```{r}
# Poker and roulette winnings from Monday to Friday:
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Define a new variable based on a selection
poker_midweek <- 
```

*** =solution
```{r}
# Poker and roulette winnings from Monday to Friday:
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Define a new variable based on a selection
poker_midweek <- poker_vector[c(2, 3, 4)]
```

*** =sct
```{r}
msg = "Do not change anything about the definition and naming of `poker_vector` and `roulette_vector`."
test_object("days_vector", undefined_msg = msg, incorrect_msg = msg)
test_object("poker_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)
test_object("roulette_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)
test_object("poker_midweek", 
            incorrect_msg = "It looks like `poker_midweek` does not contain the correct values from `poker_vector`. You can use the vector `c(2, 3, 4)` inside square brackets.")
success_msg("Well done! Continue to the next exercise to specialize in vector selection some more!");
```


--- type:NormalExercise xp:100 skills:1 key:27976b79f4
## Vector selection: the good times (3)

Selecting multiple elements of `poker_vector` with `c(2, 3, 4)` is not very convenient. Many statisticians are lazy people by nature, so they created an easier way to do this: `c(2, 3, 4)` can be abbreviated to`2:4`, which generates a vector with all natural numbers from 2 up to 4.

So, another way to find the mid-week results is `poker_vector[2:4]`. Notice how the vector `2:4` is placed between the square brackets to select element 2 up to 4.

*** =instructions
Assign to `roulette_selection_vector` the roulette results from Tuesday up to Friday; make use of `:` if it makes things easier for you.

*** =hint
Assign a selection of `roulette_vector` to `roulette_selection_vector` by placing `2:5` between square brackets.

*** =sample_code
```{r}
# Poker and roulette winnings from Monday to Friday:
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Define a new variable based on a selection
roulette_selection_vector <- 
```

*** =solution
```{r}
# Poker and roulette winnings from Monday to Friday:
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Define a new variable based on a selection
roulette_selection_vector <- roulette_vector[2:5]
```

*** =sct

```{r}
msg = "Do not change anything about the definition and naming of `poker_vector` and `roulette_vector`."
test_object("days_vector", undefined_msg = msg, incorrect_msg = msg)
test_object("poker_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)
test_object("roulette_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)
test_object("roulette_selection_vector", 
            undefined_msg = "Please make sure to define a variable `roulette_selection_vector`.",
            incorrect_msg = "It looks like `roulette_selection_vector` does not contain the correct selection from `roulette_vector`. Make sure to to use the right indexes.")
success_msg("Awesome! The colon operator is extremely useful and very often used in R programming, so remember it well. Proceed to the next exercise.")
```


--- type:NormalExercise xp:100 skills:1 key:e6c263ddee
## Vector selection: the good times (4)

Another way to tackle the previous exercise is by using the names of the vector elements (Monday, Tuesday, ...) instead of their numeric positions. For example, 

```
poker_vector["Monday"]
```

will select the first element of `poker_vector` since `"Monday"` is the name of that first element.

Just like you did in the previous exercise with numerics, you can also use the element names to select multiple elements, for example: 

```
poker_vector[c("Monday","Tuesday")]
```

*** =instructions
- Select the first three elements in `poker_vector` by using their names: `"Monday"`, `"Tuesday"` and `"Wednesday"`. Assign the result of the selection to `poker_start`.
- Calculate the average of the values in `poker_start` with the [`mean()`](http://www.rdocumentation.org/packages/base/functions/mean) function. Simply print out the result so you can inspect it.

*** =hint
- You can use `c("Monday", "Tuesday", "Wednesday")` inside square brackets to subset `poker_vector` appropriately.
- You can use `mean(poker_start)` to get the mean of the elements in `poker_start`. You do not need the mean of all poker elements, but only of the first three days.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# Poker and roulette winnings from Monday to Friday:
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Select poker results for Monday, Tuesday and Wednesday
poker_start <- 
  
# Calculate the average of the elements in poker_start

```

*** =solution
```{r}
# Poker and roulette winnings from Monday to Friday:
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Select poker results for Monday, Tuesday and Wednesday
poker_start <- poker_vector[c("Monday", "Tuesday", "Wednesday")]
  
# Calculate the average of the elements in poker_start
mean(poker_start)
```

*** =sct
```{r}
msg = "Do not change anything about the definition and naming of `poker_vector` and `roulette_vector`."
test_object("days_vector", undefined_msg = msg, incorrect_msg = msg)
test_object("poker_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)
test_object("roulette_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)
test_object("poker_start", 
            incorrect_msg = "It looks like `poker_start` does not contain the first three values of `poker_vector`. You can use `c(\"Monday\", \"Tuesday\", \"Wednesday\")` inside square brackets to do this.")
test_output_contains("mean(poker_start)", incorrect_msg = "Have you correctly calculated the average of the values in `poker_start` and printed it out? Use `mean(poker_start)`.")
success_msg("Good job! Apart from subsetting vectors by index or by name, you can also subset vectors by comparison. The next exercises will show you how!");
```


--- type:NormalExercise xp:100 skills:1 key:f0f619c901
## Selection by comparison - Step 1

By making use of comparison operators, we can approach the previous question in a more proactive way. 

The (logical) comparison operators known to R are:

- `<` for less than
- `>` for greater than
- `<=` for less than or equal to
- `>=` for greater than or equal to
- `==` for equal to each other
- `!=` not equal to each other

As seen in the previous chapter, stating `6 > 5` returns `TRUE`. The nice thing about R is that you can use these comparison operators also on vectors. For example:

```
> c(4, 5, 6) > 5
[1] FALSE FALSE TRUE
```

This command tests for every element of the vector if the condition stated by the comparison operator is `TRUE` or `FALSE`.

*** =instructions
- Check which elements in `poker_vector` are positive (i.e. > 0) and assign this to `selection_vector`. 
- Print out `selection_vector` so you can inspect it. The printout tells you whether you won (`TRUE`) or lost (`FALSE`) any money for each day.

*** =hint
In order to check for which days your poker gains are positive, R should check for each element of `poker_vector` whether it is larger than zero. `some_vector > 0` is the way to tell R what you are after.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# Poker and roulette winnings from Monday to Friday:
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Which days did you make money on poker?
selection_vector <- 
  
# Print out selection_vector

```

*** =solution
```{r}
# Poker and roulette winnings from Monday to Friday:
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Which days did you make money on poker?
selection_vector <- poker_vector > 0
  
# Print out selection_vector
selection_vector
```

*** =sct
```{r}
msg <- "Do not change anything about the definition and naming of `poker_vector` and `roulette_vector`."
test_object("days_vector", undefined_msg = msg, incorrect_msg = msg)
test_object("poker_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)
test_object("roulette_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)

test_object("selection_vector", incorrect_msg = "It looks like `selection_vector` does not contain the correct result. Remember that R uses element wise operations for vectors.")
test_output_contains("selection_vector", incorrect_msg = "Don't forget to print out `selection_vector` by writing the variable name on a new line.")
success_msg("Great!")
```


--- type:NormalExercise xp:100 skills:1 key:2754fc5cd4
## Selection by comparison - Step 2

Working with comparisons will make your data analytical life easier. Instead of selecting a subset of days to investigate yourself (like before), you can simply ask R to return only those days where you realized a positive return for poker. 

In the previous exercises you used `selection_vector <- poker_vector > 0` to find the days on which you had a positive poker return. Now, you would like to know not only the days on which you won, but also how much you won on those days. 

You can select the desired elements, by putting `selection_vector` between the square brackets that follow `poker_vector`:

```
poker_vector[selection_vector]
```

R knows what to do when you pass a logical vector in square brackets: it will only select the elements that correspond to `TRUE` in `selection_vector`.

*** =instructions
Use `selection_vector` in square brackets to assign the amounts that you won on the profitable days to the variable `poker_winning_days`.

*** =hint
Use `poker_vector[selection_vector]` to select the desired elements from `poker_vector`, and assign the result to `poker_winning_days`.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# Poker and roulette winnings from Monday to Friday:
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Which days did you make money on poker?
selection_vector <- poker_vector > 0

# Select from poker_vector these days
poker_winning_days <- 
```

*** =solution
```{r}
# Poker and roulette winnings from Monday to Friday:
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Which days did you make money on poker?
selection_vector <- poker_vector > 0

# Select from poker_vector these days
poker_winning_days <- poker_vector[selection_vector]
```

*** =sct
```{r}
msg = "Do not change anything about the definition and naming of `poker_vector` and `roulette_vector`."
test_object("days_vector", undefined_msg = msg, incorrect_msg = msg)
test_object("poker_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)
test_object("roulette_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)
test_object("selection_vector", incorrect_msg = "Don't change the way `selection_vector` is calculated.")
test_object("poker_winning_days",
            incorrect_msg =  "It looks like `poker_winning_days` does not contain the correct result. Use `poker_vector[selection_vector]`.")
success_msg("Good job! Continue to the next exercise.")
```


--- type:NormalExercise xp:100 skills:1 key:59e8dcbbd5
## Advanced selection

Just like you did for poker, you also want to know those days where you realized a positive return for roulette.

*** =instructions
- Create the variable `selection_vector`, this time to see if you made profit with roulette for different days.
- Assign the amounts that you made on the days that you ended positively for roulette to the variable `roulette_winning_days`. This vector thus contains the positive winnings of `roulette_vector`.

*** =hint
Once you've correctly calculated `selection_vector`, you can again use `roulette_vector[selection_vector]` to select the positive results from `roulette_vector`.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# Poker and roulette winnings from Monday to Friday:
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Which days did you make money on roulette?
selection_vector <-

# Select from roulette_vector these days
roulette_winning_days <- 
```

*** =solution
```{r}
# Poker and roulette winnings from Monday to Friday:
poker_vector <- c(140, -50, 20, -120, 240)
roulette_vector <- c(-24, -50, 100, -350, 10)
days_vector <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
names(poker_vector) <- days_vector
names(roulette_vector) <- days_vector

# Which days did you make money on roulette?
selection_vector <- roulette_vector > 0

# Select from roulette_vector these days
roulette_winning_days <- roulette_vector[selection_vector]
```

*** =sct
```{r}
msg = "Do not change anything about the definition and naming of `poker_vector` and `roulette_vector`."
test_object("days_vector", undefined_msg = msg, incorrect_msg = msg)
test_object("poker_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)
test_object("roulette_vector", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)
test_object("selection_vector", 
            incorrect_msg = "It looks like `selection_vector` does not contain the correct result. Use `roulette_vector > 0`.")
test_object("roulette_winning_days",
            incorrect_msg = "It looks like `roulette_winning_days` does not contain the correct result. Use `roulette_vector[selection_vector]`.")

success_msg("Great! This exercise concludes the chapter on vectors. The next chapter will introduce you to the two-dimensional version of vectors: matrices.")
```



