--- 
title_meta  : บทที่ 3
title       : Matrices
description : ในบทนี้ คุณจะได้เรียนรู้เกี่ยวกับการใช้ matrix ใน R เมื่อคุณทำแบบฝึกหัดในบทนี้ครบแล้ว คุณจะรู้จักวิธีการสร้าง matrix และใช้มันในการคำนวณได้ เราจะให้คุณฝึกใช้ matrix โดยการวิเคราะห์ตัวเลขรายได้ของภาพยนตร์เรื่อง Star Wars ขอพลังจงสถิตอยู่กับท่าน!

--- type:NormalExercise xp:100 skills:1 key:d61aeba84c
## Matrix คืออะไร?

ใน R matrix คือการเก็บข้อมูลประเภทเดียวกัน (numeric, character, หรือ logical) หลายๆตัวไว้ในรูปของตารางซึ่งประกอบไปด้วยแถวและคอลัมน์(ลองนึกภาพ vector หลายๆอันมาต่อกัน) และเนื่องจากเราสามารถระบุตำแหน่งของข้อมูลแต่ละตัวได้ด้วยเลขแถวและคอลัมน์ matrix จึงมี 2 มิติ! 

คุณสามารถสร้าง matrix ได้ด้วยการใช้ function [`matrix()`](http://www.rdocumentation.org/packages/base/functions/matrix) ลองดูโค้ดด้านล่างเป็นตัวอย่าง:

```
matrix(1:9, byrow = TRUE, nrow = 3)
```

เมื่อคุณใช้ function [`matrix()`](http://www.rdocumentation.org/packages/base/functions/matrix):

- argument ตัวแรกหมายถึงข้อมูลที่เราจะใช้เข้าไปเพื่อจัดเรียงให้เป็นแถวและคอลัมน์เพื่อสร้าง matrix ในที่นี้เราใช้ `1:9` ซึ่งมีค่าเท่ากับ `c(1, 2, 3, 4, 5, 6, 7, 8, 9)`
- argument ตัวที่สอง `byrow` เป็นการบอกว่าจะสร้าง matrix นี้โดยการเรียงข้อมูลตามแนวแถว(คือเรียงข้อมูลไล่จากซ้ายไปขวา) หรือไม่ ถ้าคุณต้องการให้เรียงตามแนวคอลัมน์(คือเรียงข้อมูลไล่จากบนลงล่าง) ให้คุณใส่ argument นี้ว่า `byrow = FALSE` แทน 
- argument ตัวที่สาม `nrow` เป็นการบอก R ว่าเราต้องการ matrix ที่มีขนาดกี่แถว

*** =instructions
ลองสร้าง matrix ที่มี 3 แถว โดยการใส่เลข 1 ถึง 9 และเรียงข้อมูลตามแนวแถว

*** =hint
ลองอ่านคำสั่งดูดีๆ คำตอบอยู่ในนั้นแล้ว!

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# Construct a matrix with 3 rows that contain the numbers 1 up to 9
```

*** =solution
```{r}
# Construct a matrix with 3 rows that contain the numbers 1 up to 9
matrix(1:9, byrow = TRUE, nrow = 3)
```

*** =sct
```{r}
function_msg <- "&#3588;&#3640;&#3603;&#3652;&#3604;&#3657;&#3626;&#3619;&#3657;&#3634;&#3591;&#32;&#109;&#97;&#116;&#114;&#105;&#120;&#32;&#3629;&#3618;&#3656;&#3634;&#3591;&#3606;&#3641;&#3585;&#3605;&#3657;&#3629;&#3591;&#3627;&#3619;&#3639;&#3629;&#3648;&#3611;&#3621;&#3656;&#3634;&#63;&#32;&#3621;&#3629;&#3591;&#3604;&#3641;&#3588;&#3635;&#3626;&#3633;&#3656;&#3591;&#3629;&#3637;&#3585;&#3588;&#3619;&#3633;&#3657;&#3591;&#32;&#3588;&#3635;&#3605;&#3629;&#3610;&#3629;&#3618;&#3641;&#3656;&#3651;&#3609;&#3609;&#3633;&#3657;&#3609;&#3649;&#3627;&#3621;&#3632;"
# คุณได้สร้าง matrix อย่างถูกต้องหรือเปล่า? ลองดูคำสั่งอีกครั้ง คำตอบอยู่ในนั้นแหละ

output_msg <- "&#3648;&#3627;&#3617;&#3639;&#3629;&#3609;&#3588;&#3640;&#3603;&#3592;&#3632;&#3617;&#3637;&#3611;&#3633;&#3597;&#3627;&#3634;&#3585;&#3633;&#3610;&#3585;&#3634;&#3619;&#3626;&#3619;&#3657;&#3634;&#3591;&#32;&#109;&#97;&#116;&#114;&#105;&#120;&#32;&#3609;&#3632;&#32;&#3621;&#3629;&#3591;&#3604;&#3641;&#3588;&#3635;&#3626;&#3633;&#3656;&#3591;&#3629;&#3637;&#3585;&#3588;&#3619;&#3633;&#3657;&#3591;&#32;&#3588;&#3635;&#3605;&#3629;&#3610;&#3629;&#3618;&#3641;&#3656;&#3651;&#3609;&#3609;&#3633;&#3657;&#3609;&#3649;&#3621;&#3657;&#3623;"
# เหมือนคุณจะมีปัญหากับการสร้าง matrix นะ ลองดูคำสั่งอีกครั้ง คำตอบอยู่ในนั้นแล้ว

success_msg <- "&#3648;&#3618;&#3637;&#3656;&#3618;&#3617;&#3617;&#3634;&#3585;&#33;&#32;&#3652;&#3611;&#3649;&#3610;&#3610;&#3613;&#3638;&#3585;&#3627;&#3633;&#3604;&#3605;&#3656;&#3629;&#3652;&#3611;&#3585;&#3633;&#3609;&#3648;&#3621;&#3618;"
# เยี่ยมมาก! ไปแบบฝึกหัดต่อไปกันเลย

test_function("matrix", c("data", "byrow", "nrow"), incorrect_msg = function_msg)
test_output_contains("matrix(1:9, byrow=TRUE, nrow=3)", incorrect_msg = output_msg)
success_msg(success_msg)
```


--- type:NormalExercise xp:100 skills:1 key:effc2fb945
## การสร้าง matrix

มาถึงของจริงกันบ้าง เราจะให้คุณได้ลองเริ่มวิเคราะห์รายได้ box office ของภาพยนตร์เรื่อง Star Wars โดยใช้ matrix ขอพลังจงสถิตอยู่กับท่าน! 

ใน editor vectors ทั้ง 3 ตัวถูกสร้างรอไว้แล้ว โดยเราให้แต่ละตัวแทนรายได้จากภาพยนตร์ Star Wars 3 ภาคแรก ข้อมูลตัวแรกใน vector แทนรายได้จาก US box office ตัวที่สองแทนรายได้จาก box office อื่นๆที่ไม่ใช่ US (แหล่งข้อมูล: Wikipedia).

ในแบบฝึกหัดนี้ ให้คุณรวมตัวเลขใน vector เหล่านี้เข้าด้วยกันเป็น vector เดียว จากนั้นสร้าง matrix จาก vector ดังกล่าว

*** =instructions
- ใช้คำสั่ง `c(new_hope, empire_strikes, return_jedi)` ในการรวม vector ทั้งสามเข้าด้วยกันเป็นหนึ่ง แล้วให้ vector ใหม่นี้มีชื่อว่า `box_office`
- ใช้ function `matrix()` ในการสร้าง matrix ที่มี 3 แถว แต่ละแถวแทนข้อมูลของภาพยนตร์แต่ละภาค โดยให้ argument ตัวแรกเป็น vector `box_office` ซึ่งเก็บข้อมูลรายได้ภาพยนตร์ทั้งหมดไว้ จากนั้่นคุณต้องกำหนดให้ `nrow = 3` และ `byrow = TRUE` ตั้งชื่อ matrix ว่า `star_wars_matrix`.

*** =hint
- คำสั่ง `box_office <- c(new_hope, empire_strikes, return_jedi)` จะทำการรวม vector เข้าด้วยกันเป็นอันเดียว ซึ่ง vector ใหม่จะมีข้อมูล 6 ตัวด้วยกัน
- ลองดูคำสั่ง `matrix(box_office, nrow = ..., by_row ...)` แล้วนำไปปรับใช้ให้ถูกต้อง

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# Box office Star Wars (in millions!)
new_hope <- c(460.998, 314.4)
empire_strikes <- c(290.475, 247.900)
return_jedi <- c(309.306, 165.8)

# Create box_office
box_office <- 

# Construct star_wars_matrix
star_wars_matrix <- 
```

*** =solution
```{r}
# Box office Star Wars (in millions!)
new_hope <- c(460.998, 314.4)
empire_strikes <- c(290.475, 247.900)
return_jedi <- c(309.306, 165.8)

# Create box_office
box_office <- c(new_hope, empire_strikes, return_jedi)

# Construct star_wars_matrix
star_wars_matrix <- matrix(box_office, nrow = 3, byrow = TRUE) 
```

*** =sct
```{r}
msg <- paste("&#3629;&#3618;&#3656;&#3634;&#3648;&#3611;&#3621;&#3637;&#3656;&#3618;&#3609;&#3649;&#3611;&#3621;&#3591;&#3605;&#3633;&#3623;&#3649;&#3611;&#3619;", "`new_hope`", "`empire_strikes`", "&#3649;&#3621;&#3632;", "`return_jedi`!")
# อย่าเปลี่ยนแปลงตัวแปร `new_hope`, `empire_strikes` และ `return_jedi`!

box_office_msg <- paste("&#3588;&#3640;&#3603;&#3652;&#3604;&#3657;&#3609;&#3635;&#3586;&#3657;&#3629;&#3617;&#3641;&#3621;&#3592;&#3634;&#3585;", "`new_hope` `empire_strikes`", "&#3649;&#3621;&#3632;", "`return_jedi`", "&#3617;&#3634;&#3619;&#3623;&#3617;&#3585;&#3633;&#3609;&#3629;&#3618;&#3656;&#3634;&#3591;&#3606;&#3641;&#3585;&#3605;&#3657;&#3629;&#3591;&#3648;&#3611;&#3655;&#3609;", "`box_office`", "&#3627;&#3619;&#3639;&#3629;&#3618;&#3633;&#3591;&#63;")
# คุณได้นำข้อมูลจาก `new_hope` `empire_strikes` และ `return_jedi` มารวมกันอย่างถูกต้องเป็น vector `box_office` หรือยัง?

matrix_msg <- paste("&#3605;&#3619;&#3623;&#3592;&#3626;&#3629;&#3610;&#3651;&#3627;&#3657;&#3649;&#3609;&#3656;&#3651;&#3592;&#3623;&#3656;&#3634;&#3588;&#3640;&#3603;&#3652;&#3604;&#3657;&#3651;&#3626;&#3656;&#32;&#97;&#114;&#103;&#117;&#109;&#101;&#110;&#116;&#32;&#3605;&#3656;&#3634;&#3591;&#3654;&#3651;&#3627;&#3657;", "`matrix()`", "&#3629;&#3618;&#3656;&#3634;&#3591;&#3606;&#3641;&#3585;&#3605;&#3657;&#3629;&#3591;", "(`box_office`, `nrow = 3`, `by_row = TRUE`)")
# ตรวจสอบให้แน่ใจว่าคุณได้ใส่ argument ต่างๆให้ function `matrix()` อย่างถูกต้อง (`box_office`, `nrow = 3`, `by_row = TRUE`)

star_wars_matrix <- paste("&#3588;&#3640;&#3603;&#3652;&#3604;&#3657;&#3585;&#3635;&#3627;&#3609;&#3604;&#3651;&#3627;&#3657;", "`star_wars_matrix`", "&#3617;&#3637;&#3588;&#3656;&#3634;&#3648;&#3607;&#3656;&#3634;&#3585;&#3633;&#3610;&#32;&#109;&#97;&#116;&#114;&#105;&#120;&#32;&#3607;&#3637;&#3656;&#3588;&#3640;&#3603;&#3626;&#3619;&#3657;&#3634;&#3591;&#3627;&#3619;&#3639;&#3629;&#3648;&#3611;&#3621;&#3656;&#3634;&#63;")
# คุณได้กำหนดให้ `star_wars_matrix` มีค่าเท่ากับ matrix ที่คุณสร้างหรือเปล่า?

success_msg <- "&#3586;&#3629;&#3614;&#3621;&#3633;&#3591;&#3592;&#3591;&#3626;&#3606;&#3636;&#3605;&#3629;&#3618;&#3641;&#3656;&#3585;&#3633;&#3610;&#3607;&#3656;&#3634;&#3609;&#33;&#32;&#3648;&#3619;&#3634;&#3652;&#3611;&#3604;&#3641;&#3649;&#3610;&#3610;&#3613;&#3638;&#3585;&#3627;&#3633;&#3604;&#3605;&#3656;&#3629;&#3652;&#3611;&#3585;&#3633;&#3609;&#3648;&#3621;&#3618;"
# ขอพลังจงสถิตอยู่กับท่าน! เราไปดูแบบฝึกหัดต่อไปกันเลย

test_object("new_hope", undefined_msg = msg, incorrect_msg = msg)
test_object("empire_strikes", undefined_msg = msg, incorrect_msg = msg)
test_object("return_jedi", undefined_msg = msg, incorrect_msg = msg)

test_object("box_office", incorrect_msg = box_office_msg)
test_function("matrix", c("data", "nrow", "byrow"), incorrect_msg = matrix_msg)
test_object("star_wars_matrix", incorrect_msg = star_wars_matrix)
success_msg(success_msg)
```


--- type:NormalExercise xp:100 skills:1 key:f734e8bf74
## การตั้งชื่อให้กับข้อมูลใน matrix

การตั้งชื่อของภาพยนตร์ให้กับแต่ละแถวใน matrix `star_wars_matrix` จะช่วยให้คุณอ่านข้อมูลได้ง่ายขึ้น แถมยังมีประโยชน์เวลาคุณจะเรียกใช้ข้อมูลอีกด้วย! 

คล้ายๆกับตอนที่เราตั้งชื่อให้กับข้อมูลใน vector คุณสามารถตั้งชื่อให้กับแถวและคอลัมน์ของ matrix ได้

```
rownames(my_matrix) <- row_names_vector
colnames(my_matrix) <- col_names_vector
```

เราเตรียม vector สองอันนี้ไว้ให้คุณ: `region` และ `titles` ให้คุณใช้ vector ทั้งสองในการตั้งชื่อให้กับคอลัมน์และแถวของ `star_wars_matrix` (เรียงตามลำดับ)

*** =instructions
- ใช้คำสั่ง `colnames()` ในการตั้งชื่อแต่ละคอลัมน์ของ `star_wars_matrix` ด้วย `region` vector
- ใช้คำสั่ง `rownames()` ในการตั้งชื่อแต่ละแถวของ `star_wars_matrix` ด้วย `titles` vector
- ลองสั่งแสดงค่า `star_wars_matrix` แล้วดูว่าผลลัพธ์เป็นอย่างไร

*** =hint
คุณสามารถใช้คำสั่ง `colnames(star_wars_matrix) <- region` ในการตั้งชื่อให้กับแต่ละคอลัมน์ของ `star_wars_matrix` ลองปรับเปลี่ยนคำสั่งนี้และนำไปตั้งชื่อให้กับแต่ละแถวดูบ้าง

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# Box office Star Wars (in millions!)
new_hope <- c(460.998, 314.4)
empire_strikes <- c(290.475, 247.900)
return_jedi <- c(309.306, 165.8)

# Construct matrix
star_wars_matrix <- matrix(c(new_hope, empire_strikes, return_jedi), nrow = 3, byrow = TRUE)

# Vectors region and titles, used for naming
region <- c("US", "non-US")
titles <- c("A New Hope", "The Empire Strikes Back", "Return of the Jedi")

# Name the columns with region


# Name the rows with titles


# Print out star_wars_matrix
```

*** =solution
```{r}
# Box office Star Wars (in millions!)
new_hope <- c(460.998, 314.4)
empire_strikes <- c(290.475, 247.900)
return_jedi <- c(309.306, 165.8)

# Construct matrix
star_wars_matrix <- matrix(c(new_hope, empire_strikes, return_jedi), nrow = 3, byrow = TRUE)

# Vectors region and titles, used for naming
region <- c("US", "non-US")
titles <- c("A New Hope", "The Empire Strikes Back", "Return of the Jedi")

# Name the columns with region
colnames(star_wars_matrix) <- region

# Name the rows with titles
rownames(star_wars_matrix) <- titles

# Print out star_wars_matrix
star_wars_matrix
```

*** =sct
```{r}
msg <- paste("&#3629;&#3618;&#3656;&#3634;&#3648;&#3611;&#3621;&#3637;&#3656;&#3618;&#3609;&#3649;&#3611;&#3621;&#3591;&#3605;&#3633;&#3623;&#3649;&#3611;&#3619;", "`new_hope`", "`empire_strikes`", "&#3649;&#3621;&#3632;", "`return_jedi`!")
# อย่าเปลี่ยนแปลงตัวแปร `new_hope`, `empire_strikes` และ `return_jedi`!

test_object("new_hope", undefined_msg = msg, incorrect_msg = msg)
test_object("empire_strikes", undefined_msg = msg, incorrect_msg = msg)
test_object("return_jedi", undefined_msg = msg, incorrect_msg = msg)

msg <- paste("&#3629;&#3618;&#3656;&#3634;&#3648;&#3611;&#3621;&#3637;&#3656;&#3618;&#3609;&#3649;&#3611;&#3621;&#3591;&#3588;&#3656;&#3634;&#3616;&#3634;&#3618;&#3651;&#3609;", "`star_wars_matrix`", "&#3651;&#3627;&#3657;&#3648;&#3611;&#3621;&#3637;&#3656;&#3618;&#3609;&#3649;&#3588;&#3656;&#3594;&#3639;&#3656;&#3629;&#3586;&#3629;&#3591;&#3649;&#3606;&#3623;&#3649;&#3621;&#3632;&#3588;&#3629;&#3621;&#3633;&#3617;&#3609;&#3660;&#3648;&#3607;&#3656;&#3634;&#3609;&#3633;&#3657;&#3609;")
# อย่าเปลี่ยนแปลงค่าภายใน `star_wars_matrix` ให้เปลี่ยนแค่ชื่อของแถวและคอลัมน์เท่านั้น

test_object("star_wars_matrix", incorrect_msg = msg)
msg <- paste("&#3629;&#3618;&#3656;&#3634;&#3648;&#3611;&#3621;&#3637;&#3656;&#3618;&#3609;&#3649;&#3611;&#3621;&#3591;&#3605;&#3633;&#3623;&#3649;&#3611;&#3619;", "`region`", "&#3649;&#3621;&#3632;", "`titles`")
# อย่าเปลี่ยนแปลงตัวแปร `region` และ `titles`

test_object("region", undefined_msg = msg, incorrect_msg = msg)
test_object("titles", undefined_msg = msg, incorrect_msg = msg)

object_msg <- paste("&#3588;&#3640;&#3603;&#3652;&#3604;&#3657;&#3605;&#3633;&#3657;&#3591;&#3594;&#3639;&#3656;&#3629;&#3651;&#3627;&#3657;&#3649;&#3606;&#3623;&#3649;&#3621;&#3632;&#3588;&#3629;&#3621;&#3633;&#3617;&#3609;&#3660;&#3586;&#3629;&#3591;", "`star_wars_matrix`", "&#3629;&#3618;&#3656;&#3634;&#3591;&#3606;&#3641;&#3585;&#3605;&#3657;&#3629;&#3591;&#3627;&#3619;&#3639;&#3629;&#3648;&#3611;&#3621;&#3656;&#3634;&#63;&#32;&#3651;&#3594;&#3657;&#3588;&#3635;&#3626;&#3633;&#3656;&#3591;", "`colnames(star_wars_matrix) <- region`", "&#3651;&#3609;&#3585;&#3634;&#3619;&#3605;&#3633;&#3657;&#3591;&#3594;&#3639;&#3656;&#3629;&#3588;&#3629;&#3621;&#3633;&#3617;&#3609;&#3660;&#32;&#3649;&#3621;&#3632;&#3651;&#3594;&#3657;&#3588;&#3635;&#3626;&#3633;&#3656;&#3591;&#3588;&#3621;&#3657;&#3634;&#3618;&#3654;&#3585;&#3633;&#3609;&#3651;&#3609;&#3585;&#3634;&#3619;&#3605;&#3633;&#3657;&#3591;&#3594;&#3639;&#3656;&#3629;&#3651;&#3627;&#3657;&#3649;&#3606;&#3623;")
# คุณได้ตั้งชื่อให้แถวและคอลัมน์ของ `star_wars_matrix` อย่างถูกต้องหรือเปล่า? ใช้คำสั่ง `colnames(star_wars_matrix) <- region` ในการตั้งชื่อคอลัมน์ และใช้คำสั่งคล้ายๆกันในการตั้งชื่อให้แถว

output_msg <- paste("&#3629;&#3618;&#3656;&#3634;&#3621;&#3639;&#3617;&#3626;&#3633;&#3656;&#3591;&#3651;&#3627;&#3657;&#32;&#82;&#32;&#3649;&#3626;&#3604;&#3591;&#3588;&#3656;&#3634;", "`star_wars_matrix`", "&#3627;&#3621;&#3633;&#3591;&#3592;&#3634;&#3585;&#3588;&#3640;&#3603;&#3652;&#3604;&#3657;&#3605;&#3633;&#3657;&#3591;&#3594;&#3639;&#3656;&#3629;&#3651;&#3627;&#3657;&#3585;&#3633;&#3610;&#3649;&#3606;&#3623;&#3649;&#3621;&#3632;&#3588;&#3629;&#3621;&#3633;&#3617;&#3609;&#3660;&#3649;&#3621;&#3657;&#3623;")
# อย่าลืมสั่งให้ R แสดงค่า `star_wars_matrix` หลังจากคุณได้ตั้งชื่อให้กับแถวและคอลัมน์แล้ว

success_msg <- "&#3607;&#3635;&#3652;&#3604;&#3657;&#3604;&#3637;&#3617;&#3634;&#3585;&#33;&#32;&#3588;&#3640;&#3603;&#3585;&#3635;&#3621;&#3633;&#3591;&#3629;&#3618;&#3641;&#3656;&#3610;&#3609;&#3648;&#3626;&#3657;&#3609;&#3607;&#3634;&#3591;&#3649;&#3627;&#3656;&#3591;&#3585;&#3634;&#3619;&#3648;&#3611;&#3655;&#3609;&#3612;&#3641;&#3657;&#3648;&#3594;&#3637;&#3656;&#3618;&#3623;&#3594;&#3634;&#3597;&#32;&#82;&#32;&#3649;&#3621;&#3657;&#3623;&#33;&#32;&#3652;&#3611;&#3649;&#3610;&#3610;&#3613;&#3638;&#3585;&#3627;&#3633;&#3604;&#3605;&#3656;&#3629;&#3652;&#3611;&#3585;&#3633;&#3609;&#3648;&#3621;&#3618;"
# ทำได้ดีมาก! คุณกำลังอยู่บนเส้นทางแห่งการเป็นผู้เชี่ยวชาญ R แล้ว! ไปแบบฝึกหัดต่อไปกันเลย

test_object("star_wars_matrix", eq_condition = "equal", incorrect_msg = object_msg)
test_output_contains("star_wars_matrix", incorrect_msg = output_msg)
success_msg(success_msg)
```


--- type:NormalExercise xp:100 skills:1 key:3fd7499a12
## การคำนวณรายได้จาก box office ทั่วโลก

The single most important thing for a movie in order to become an instant legend in Tinseltown is its worldwide box office figures. 

To calculate the total box office revenue for the three Star Wars movies, you have to take the sum of the US revenue column and the non-US revenue column. 

In R, the function [`rowSums()`](http://www.rdocumentation.org/packages/base/functions/colSums) conveniently calculates the totals for each row of a matrix. This function creates a new vector:

```
rowSums(my_matrix)
```

*** =instructions
Calculate the worldwide box office figures for the three movies and put these in the vector named `worldwide_vector`.

*** =hint
`rowSums(star_wars_matrix)` will calculate the sum of every row, so the total box office for each movie.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# Construct star_wars_matrix
box_office <- c(460.998, 314.4, 290.475, 247.900, 309.306, 165.8)
star_wars_matrix <- matrix(box_office, nrow = 3, byrow = TRUE,
                           dimnames = list(c("A New Hope", "The Empire Strikes Back", "Return of the Jedi"), 
                                           c("US", "non-US")))

# Calculate worldwide box office figures
worldwide_vector <- 
```

*** =solution
```{r}
# Construct star_wars_matrix
box_office <- c(460.998, 314.4, 290.475, 247.900, 309.306, 165.8)
star_wars_matrix <- matrix(box_office, nrow = 3, byrow = TRUE,
                           dimnames = list(c("A New Hope", "The Empire Strikes Back", "Return of the Jedi"), 
                                           c("US", "non-US")))

# Calculate worldwide box office figures
worldwide_vector <- rowSums(star_wars_matrix)
```

*** =sct
```{r}
msg <- paste("&#3629;&#3618;&#3656;&#3634;&#3648;&#3611;&#3621;&#3637;&#3656;&#3618;&#3609;&#3649;&#3611;&#3621;&#3591;&#3605;&#3633;&#3623;&#3649;&#3611;&#3619;", "`box_office_all`", "&#3649;&#3621;&#3632;", "`star_wars_marix`!")
# อย่าเปลี่ยนแปลงตัวแปร `box_office_all` และ `star_wars_marix`!

msg <- "Do not change anything about the preset variables `box_office_all` and `star_wars_marix`!"
test_object("box_office", undefined_msg = msg, incorrect_msg = msg)
test_object("star_wars_matrix", undefined_msg = msg, incorrect_msg = msg)
test_object("worldwide_vector", incorrect_msg = "Call `rowSums()` on `star_wars_matrix` and store the result in `worldwide_vector`.")
success_msg("Well done! Continue to the next exercise.")
```


--- type:NormalExercise xp:100 skills:1 key:86b87a8545
## Adding a column for the Worldwide box office 

In the previous exercise you calculated the vector that contained the worldwide box office receipt for each of the three Star Wars movies. However, this vector is not yet part of `star_wars_matrix`.

You can add a column or multiple columns to a matrix with the [`cbind()`](http://www.rdocumentation.org/packages/base/functions/cbind) function, which merges matrices and/or vectors together by column. For example: 

```
big_matrix <- cbind(matrix1, matrix2, vector1 ...)
```

*** =instructions
Add `worldwide_vector` as a new column to the `star_wars_matrix` and assign the result to `all_wars_matrix`. Use the [`cbind()`](http://www.rdocumentation.org/packages/base/functions/cbind) function.

*** =hint
In this exercise, you should pass two variables to `cbind()`: `star_wars_matrix` and `worldwide_vector`, in this order. Assign the result to `all_wars_matrix`.

*** =pre_exercise_code
```{r}
# no pec
```

*** =sample_code
```{r}
# Construct star_wars_matrix
box_office <- c(460.998, 314.4, 290.475, 247.900, 309.306, 165.8)
star_wars_matrix <- matrix(box_office, nrow = 3, byrow = TRUE,
                           dimnames = list(c("A New Hope", "The Empire Strikes Back", "Return of the Jedi"), 
                                           c("US", "non-US")))

# The worldwide box office figures
worldwide_vector <- rowSums(star_wars_matrix)

# Bind the new variable worldwide_vector as a column to star_wars_matrix
all_wars_matrix <- 
```

*** =solution
```{r}
# Construct star_wars_matrix
box_office <- c(460.998, 314.4, 290.475, 247.900, 309.306, 165.8)
star_wars_matrix <- matrix(box_office, nrow = 3, byrow = TRUE,
                           dimnames = list(c("A New Hope", "The Empire Strikes Back", "Return of the Jedi"), 
                                           c("US", "non-US")))

# The worldwide box office figures
worldwide_vector <- rowSums(star_wars_matrix)

# Bind the new variable worldwide_vector as a column to star_wars_matrix
all_wars_matrix <- cbind(star_wars_matrix, worldwide_vector)
```

*** =sct
```{r}
msg <- paste("&#3629;&#3618;&#3656;&#3634;&#3648;&#3611;&#3621;&#3637;&#3656;&#3618;&#3609;&#3649;&#3611;&#3621;&#3591;&#3605;&#3633;&#3623;&#3649;&#3611;&#3619;", "`box_office_all`", "&#3649;&#3621;&#3632;", "`star_wars_marix`!")
# อย่าเปลี่ยนแปลงตัวแปร `box_office_all` และ `star_wars_marix`!

test_object("box_office", undefined_msg = msg, incorrect_msg = msg)
test_object("star_wars_matrix", undefined_msg = msg, incorrect_msg = msg)
test_object("worldwide_vector",
            incorrect_msg = "Store the result of `rowSums(star_wars_matrix)` in `worldwide_vector`.")

msg <- "Have you correctly used `cbind()` to add `worldwide_vector` to `star_wars_matrix`? You should pass `star_wars_matrix` and `world_wide_vector` to `cbind()`, in this order. The resulting matrix, `all_wars_matrix`, should consist of three rows and three columns."
test_object("all_wars_matrix", incorrect_msg = msg)
success_msg("Nice job! After adding column to a matrix, the logical next step is adding rows. Learn how in the next exercise.");
```


--- type:NormalExercise xp:100 skills:1 key:bcadb29139
## Adding a row

Just like every action has a reaction, every [`cbind()`](http://www.rdocumentation.org/packages/base/functions/cbind) has an [`rbind()`](http://www.rdocumentation.org/packages/base/functions/cbind). (We admit, we are pretty bad with metaphors.) 

Your R workspace, where all variables you defined 'live' ([check out what a workspace is](http://www.statmethods.net/interface/workspace.html)), has already been initialized and contains two matrices:

- `star_wars_matrix` that we have used all along, with data on the first trilogy,
- `star_wars_matrix2`, with similar data for the second trilogy. 

Type the name of these matrices in the console and hit Enter if you want to have a closer look. If you want to check out the contents of the workspace, you can type `ls()` in the console.

*** =instructions
Use `rbind()` to paste together `star_wars_matrix` and `star_wars_matrix2`, in this order. Assign the resulting matrix to `all_wars_matrix`.

*** =hint
Bind the two matrices together like this:
```
rbind(matrix1, matrix2)
```
Assign the result to `all_wars_matrix`.


*** =pre_exercise_code
```{r}
# Construct matrix
box_office_all <- c(461, 314.4, 290.5, 247.9, 309.3, 165.8)
movie_names <- c("A New Hope","The Empire Strikes Back","Return of the Jedi")
col_titles <- c("US","non-US")
star_wars_matrix <- matrix(box_office_all, nrow = 3, byrow = TRUE, dimnames = list(movie_names, col_titles))

# Construct matrix2
box_office_all2 <- c(474.5, 552.5, 310.7, 338.7, 380.3, 468.5)
movie_names2 <- c("The Phantom Menace", "Attack of the Clones", "Revenge of the Sith")
star_wars_matrix2 <- matrix(box_office_all2, nrow=3, byrow = TRUE, dimnames = list(movie_names2, col_titles))

# remove all except all_wars_matrix
rm(box_office_all)
rm(movie_names)
rm(col_titles)
rm(box_office_all2)
rm(movie_names2)
```

*** =sample_code
```{r}
# star_wars_matrix and star_wars_matrix2 are available in your workspace
star_wars_matrix  
star_wars_matrix2 

# Combine both Star Wars trilogies in one matrix
all_wars_matrix <- 
```

*** =solution
```{r}
# star_wars_matrix and star_wars_matrix2 are available in your workspace
star_wars_matrix  
star_wars_matrix2 

# Combine both Star Wars trilogies in one matrix
all_wars_matrix <- rbind(star_wars_matrix, star_wars_matrix2)
```

*** =sct
```{r}
msg = "Do not override the variables that have been defined for you in the workspace (`star_wars_matrix` and `star_wars_matrix2`)."
test_object("star_wars_matrix", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)
test_object("star_wars_matrix2", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)
test_object("all_wars_matrix", incorrect_msg = "Did you use the `rbind()` correctly to create `all_wars_matrix()`? `rbind()` should take two arguments: `star_wars_matrix` and `star_wars_matrix2`, in this order.")

success_msg("Wonderful! Continue with the next exercise and see how you can combine the results of the `rbind()` function with the `colSums()` function!")
```


--- type:NormalExercise xp:100 skills:1 key:1bfe5ae096
## The total box office revenue for the entire saga

Just like every [`cbind()`](http://www.rdocumentation.org/packages/base/functions/cbind) has a [`rbind()`](http://www.rdocumentation.org/packages/base/functions/cbind), every [`colSums()`](http://www.rdocumentation.org/packages/base/functions/colSums) has a [`rowSums()`](http://www.rdocumentation.org/packages/base/functions/colSums). Your R workspace already contains the `all_wars_matrix` that you constructed in the previous exercise; type `all_wars_matrix` to have another look. Let's now calculate the total box office revenue for the entire saga.

*** =instructions
- Calculate the total revenue for the US and the non-US region and assign `total_revenue_vector`. You can use the [`colSums()`](http://www.rdocumentation.org/packages/base/functions/colSums) function.
- Print out `total_revenue_vector` to have a look at the results.

*** =hint
You should use the [`colSums()`](http://www.rdocumentation.org/packages/base/functions/colSums) function with `star_wars_matrix` as the argument to find the total box office per region.

*** =pre_exercise_code
```{r}
load(url("http://s3.amazonaws.com/assets.datacamp.com/course/intro_to_r/all_wars_matrix.RData"))
```

*** =sample_code
```{r}
# all_wars_matrix is available in your workspace
all_wars_matrix

# Total revenue for US and non-US
total_revenue_vector <- 
  
# Print out total_revenue_vector
```

*** =solution
```{r}
# all_wars_matrix is available in your workspace
all_wars_matrix

# Total revenue for US and non-US
total_revenue_vector <- colSums(all_wars_matrix)

# Print out total_revenue_vector
total_revenue_vector
```

*** =sct
```{r}
msg = "Do not change the contents of `all_wars_matrix`; it was created for you in the workspace."
test_object("all_wars_matrix", eq_condition = "equal", undefined_msg = msg, incorrect_msg = msg)
test_function("colSums", "x", incorrect_msg = "Did you use the `colSums()` function on the all_wars_matrix?")
test_object("total_revenue_vector",
            incorrect_msg = "Have you correctly assigned the result of `colSums(all_wars_matrix)` to `total_revenue_vector`?")
test_output_contains("total_revenue_vector", incorrect_msg = "Don't forget to print out `total_revenue_vector`!")
success_msg("Bellissimo! Head over to the next exercise to learn matrix subsetting.")
```


--- type:NormalExercise xp:100 skills:1 key:41d9d69713
## Selection of matrix elements

Similar to vectors, you can use the square brackets `[ ]` to select one or multiple elements from a matrix. Whereas vectors have one dimension, matrices have two dimensions. You should therefore use a comma to separate that what to select from the rows from that what you want to select from the columns. For example: 

- `my_matrix[1,2]` selects the element at the first row and second column.
- `my_matrix[1:3,2:4]` results in a matrix with the data on the rows 1, 2, 3 and columns 2, 3, 4.

If you want to select all elements of a row or a column, no number is needed before or after the comma, respectively:

- `my_matrix[,1]` selects all elements of the first column.
- `my_matrix[1,]` selects all elements of the first row.

Back to Star Wars with this newly acquired knowledge! As in the previous exercise, `all_wars_matrix` is already available in your workspace.

*** =instructions
- Select the non-US revenue for all movies (the entire second column of `all_wars_matrix`), store the result as `non_us_all`.
- Use `mean()` on `non_us_all` to calculate the average non-US revenue for all movies. Simply print out the result.
- This time, select the non-US revenue for the first two movies in `all_wars_matrix`. Store the result as `non_us_some`.
- Use `mean()` again to print out the average of the values in `non_us_some`.

*** =hint
You can select the entire second column of a matrix `my_matrix` with `my_matrix[,2]`.

*** =pre_exercise_code
```{r}
load(url("http://s3.amazonaws.com/assets.datacamp.com/course/intro_to_r/all_wars_matrix.RData"))
```

*** =sample_code
```{r}
# all_wars_matrix is available in your workspace
all_wars_matrix

# Select the non-US revenue for all movies
non_us_all <- 
  
# Average non-US revenue

  
# Select the non-US revenue for first two movies
non_us_some <- 
  
# Average non-US revenue for first two movies

```

*** =solution
```{r}
# all_wars_matrix is available in your workspace
all_wars_matrix

# Select the non-US revenue for all movies
non_us_all <- all_wars_matrix[,2]
  
# Average non-US revenue
mean(non_us_all)
  
# Select the non-US revenue for first two movies
non_us_some <- all_wars_matrix[1:2,2]
  
# Average non-US revenue for first two movies
mean(non_us_some)
```

*** =sct
```{r}
msg = "Do not change the contents of `all_wars_matrix`; this matrix has already been created for you in the workspace."
test_object("all_wars_matrix", undefined_msg = msg, incorrect_msg = msg)

test_object("non_us_all",
            incorrect_msg = "Did you assign to `non_us_all` the entire second column of `all_wars_matrix`? You can use `[, 2]` to do this!")
test_output_contains("mean(non_us_all)",
                     incorrect_msg = "Have you calculated the average of the values in `non_us_all` by calling `mean(non_us_all)`? Simply print out the result.")
test_object("non_us_some",
            incorrect_msg = "Did you assign to `non_us_some` the non-US revenue for the first two movies? You can use `[1:2,2]` to do this!")
test_output_contains("mean(non_us_some)",
                     incorrect_msg = "Have you calculated the average of the values in `non_us_some` by calling `mean(non_us_some)`? Simply print out the result.")
success_msg("Nice one! Continue to the next exercise.")
```


--- type:NormalExercise xp:100 skills:1 key:c81c656f06
## A little arithmetic with matrices

Similar to what you have learned with vectors, the standard operators like `+`, `-`, `/`, `*`, etc. work in an element-wise way on matrices in R. 

For example, `2 * my_matrix` multiplies each element of `my_matrix` by two.

As a newly-hired data analyst for Lucasfilm, it is your job is to find out how many visitors went to each movie for each geographical area. You already have the total revenue figures in `all_wars_matrix`. Assume that the price of a ticket was 5 dollars. Simply dividing the box office numbers by this ticket price gives you the number of visitors.

*** =instructions
- Divide `all_wars_matrix` by 5, giving you the number of visitors in millions. Assign the resulting matrix to `visitors`.
- Print out `visitors` so you can have a look.

*** =hint
The number of visitors is equal to `all_wars_matrix` divided by 5.

*** =pre_exercise_code
```{r}
load(url("http://s3.amazonaws.com/assets.datacamp.com/course/intro_to_r/all_wars_matrix.RData"))
``` 

*** =sample_code
```{r}
# all_wars_matrix is available in your workspace
all_wars_matrix

# Estimate the visitors
visitors <- 
  
# Print the estimate to the console

```

*** =solution
```{r}
# all_wars_matrix is available in your workspace
all_wars_matrix

# Estimate the visitors
visitors <- all_wars_matrix / 5

# Print the estimate to the console
visitors
```

*** =sct
```{r}
msg = "Do not change the contents of `all_wars_matrix`; this matrix has already been created for you in the workspace."
test_object("all_wars_matrix", undefined_msg = msg, incorrect_msg = msg)

test_object("visitors",
            incorrect_msg = "It looks like `visitors` is not correct. Simply divide `all_wars_matrix` by 5 and store the resulting matrix as `visitors`.")
test_output_contains("visitors", incorrect_msg = "Don't forget to also print out `visitors` so you can have a look.")
success_msg("Great! What do these results tell you? A staggering 92 million people went to see A New Hope in US theaters! Continue to the next exercise.")
```


--- type:NormalExercise xp:100 skills:1 key:1e0b39d6e9
## A little arithmetic with matrices (2) 

Just like `2 * my_matrix` multiplied every element of `my_matrix` by two, `my_matrix1 * my_matrix2` creates a matrix where each element is the product of the corresponding elements in `my_matrix1` and `my_matrix2`. 

After looking at the result of the previous exercise, big boss Lucas points out that the ticket prices went up over time. He asks to redo the analysis based on the prices you can find in `ticket_prices_matrix` (source: imagination).

_Those who are familiar with matrices should note that this is not the standard matrix multiplication for which you should use `%*%` in R._

*** =instructions
- Divide `all_wars_matrix` by `ticket_prices_matrix` to get the estimated number of US and non-US visitors for the six movies. Assign the result to `visitors`.
- From the `visitors` matrix, select the entire first column, representing the number of visitors in the US. Store this selection as `us_visitors`.
- Calculate the average number of US visitors; print out the result.

*** =hint
- You can use the function `mean()` to calculate the average of the inputs to the function.
- To get the number of visitors in the US, select the first column from `visitors` using `visitors[ ,1]`.

*** =pre_exercise_code
```{r}
load(url("http://s3.amazonaws.com/assets.datacamp.com/course/intro_to_r/all_wars_matrix.RData"))
movie_names <- c("A New Hope","The Empire Strikes Back","Return of the Jedi", "The Phantom Menace", "Attack of the Clones", "Revenge of the Sith")
col_titles <- c("US","non-US")
ticket_prices_matrix <- matrix(c(5, 5, 6, 6, 7, 7, 4, 4, 4.5, 4.5, 4.9, 4.9), nrow = 6, byrow = TRUE, dimnames = list(movie_names,col_titles))
```

*** =sample_code
```{r}
# all_wars_matrix and ticket_prices_matrix are available in your workspace
all_wars_matrix
ticket_prices_matrix

# Estimated number of visitors
visitors <- 

# US visitors
us_visitors <- 

# Average number of US visitors

```

*** =solution
```{r}
# all_wars_matrix and ticket_prices_matrix are available in your workspace
all_wars_matrix
ticket_prices_matrix

# Estimated number of visitors
visitors <- all_wars_matrix / ticket_prices_matrix

# US visitors
us_visitors <- visitors[ ,1]

# Average number of US visitors
mean(us_visitors)
```

*** =sct
```{r}
msg <- "Do not change the contents of `all_wars_matrix`; this matrix has already been created for you in the workspace."
test_object("all_wars_matrix", undefined_msg = msg, incorrect_msg = msg)
test_object("ticket_prices_matrix", undefined_msg = msg, incorrect_msg = msg)

test_object("visitors",
            incorrect_msg = "Have you correctly created the `visitors` matrix? You should divide `all_wars_matrix` by `ticket_prices_matrix` to get there.")
test_object("us_visitors", incorrect_msg = "To created `us_visitors`, you should correctly select the entire first column from `visitors`. You can use `[,1]` for this!")
test_output_contains("mean(us_visitors)", incorrect_msg = "Once you have created `us_visitors`, you can use `mean()` to calculate the average number of visitors in the US. Make sure to print out the result.")

success_msg("It's a fact: the R force is with you! This exercise concludes the chapter on matrices. Next stop on your journey through the R language: factors.")
```


